# Accident/Incident Statistics {#sec-accident-incident-statistics}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values 

On this page is presented an overview of the **main statistics on accidents and incidents occurring in the aviation domain**, 
as well as the related fatalities 
and injuries. An overview of accidents related to ATM/ANS is also provided.
Each of the sections hereafter looks at a different aviation segment within EASA Member States. 

An *accident* constitutes an occurrence associated with the operation of an aircraft 
in which a person is fatally or seriously injured and/or the aircraft sustains
structural damage or is missing.

An *incident* is an occurrence other than an accident, which affects the safety of the aircraft.

The statistics presented hereafter are derived from EASA Annual Safety Review, 
available on [EASA website](https://www.easa.europa.eu/en/document-library/general-publications?publication_type%5b144%5d=144&page=1).
Please refer to these documents for further details and analysis on the data provided.

### Commercial air transport

In @tbl-statistics-commercial-aircraft are presented statistics on the numbers of accidents, incidents, injuries 
and fatalities for commercial air transport, which includes airline and air-taxi
passenger and cargo operations of EASA Air Operator Certificate (AOC) holders 
with aircraft of a maximum take-off mass above 5,700 kg.

```{r}
#| label: tbl-statistics-commercial-aircraft
#| tbl-cap: Accident/incident statistics - commercial air transport

read_xlsx("data/accident_statistics.xlsx",
          sheet = "cat") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    fatal = "Fatal accidents",
    nonfatal = "Non-fatal accidents",
    incidents = "Incidetns",
    fatalities = "Fatalities",
    injuries = "Injuries"
  ) %>% 
  cols_align(
    align = "center",
    columns = fatal:injuries
  )
```

### Non-commercial complex business

In this section are exposed the numbers related to the safety performance of 
complex aircraft operating non-commercial flights that are registered in EASA Member States.

```{r}
#| label: tbl-statistics-non-com-business
#| tbl-cap: Accident/incident statistics - non-commercial business aircraft

read_xlsx("data/accident_statistics.xlsx",
          sheet = "ncc") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    fatal = "Fatal accidents",
    nonfatal = "Non-fatal accidents",
    incidents = "Incidetns",
    fatalities = "Fatalities",
    injuries = "Injuries"
  ) %>% 
  cols_align(
    align = "center",
    columns = fatal:injuries
  )
```

### Specialised operations

This section covers the statistics for specialised operations (SPO) involving 
aircraft of all mass categories registered in EASA Member States.

```{r}
#| label: tbl-statistics-specialops
#| tbl-cap: Accident/incident statistics - special operations

read_xlsx("data/accident_statistics.xlsx",
          sheet = "specialops") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    fatal = "Fatal accidents",
    nonfatal = "Non-fatal accidents",
    incidents = "Incidetns",
    fatalities = "Fatalities",
    injuries = "Injuries"
  ) %>% 
  cols_align(
    align = "center",
    columns = fatal:injuries
  )
```

### Non-commercial aircraft

@tbl-statistics-noncommercial focuses on non-commercial operations involving aircraft with a 
maximum take-off mass below 5,700 kg.

```{r}
#| label: tbl-statistics-noncommercial
#| tbl-cap: Accident/incident statistics - non-commercial aircraft

read_xlsx("data/accident_statistics.xlsx",
          sheet = "noncom") %>% 
  gt() %>% 
      cols_label(
    year = "Year",
    fatal = "Fatal accidents",
    nonfatal = "Non-fatal accidents",
    incidents = "Incidetns",
    fatalities = "Fatalities",
    injuries = "Injuries"
) %>% 
      cols_align(
        align = "center",
        columns = fatal:injuries
      )  
```

### ATM statistics

Below is presented an overview of accidents and incidents statistics related to 
the provision of air traffic management or air navigation services in EASA Member States. 

Accidents and incidents considered in this section are those related to the 
provision of services by ATM/ANS providers. These represent occurrences to which
the ATM system may or may not have contributed but may have played a role in
preventing or mitigating similar occurrences in the future.

The numbers presented hereafter consider occurrences that involve at least one 
commercial aircraft (e.g. fixed-wing aircraft with a maximum take-off mass of at
least 2,250 kg, small (CS-27) or large (CS-29) helicopters) and that occurred 
in an EASA Member State.

```{r}
#| label: tbl-statistics-atm
#| tbl-cap: Accident/incident statistics - ATM/ANS

read_xlsx("data/accident_statistics.xlsx",
          sheet = "atm") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    fatal = "Fatal accidents",
    nonfatal = "Non-fatal accidents",
    incidents = "Incidetns",
    fatalities_rel = "ATM-related",
    fatalities_contr = "ATM contribution",
    injuries_rel = "ATM-related",
    injuries_contr = "ATM contribution"
  ) %>% 
  cols_align(
    align = "center",
    columns = fatal:injuries_contr
  ) %>% 
  tab_spanner(
    label = "Fatalities",
    columns = c(fatalities_rel, fatalities_contr)
  ) %>% 
  tab_spanner(
    label = "Injuries",
    columns = c(injuries_rel, injuries_contr)
  )
```

## When to use the input?

This input is recommended to be used in the instances where the safety is an
important factor, namely in terms of accident reduction.

[@ectrl:incident:report] 

[@iata:safety:report]


## References

