# Airport classification {#sec-airport-classification}

```{r}
library(readxl)
library(dplyr)
library(gt)
library(tibble)
```

## EUROCONTROL recommended values

This input represents the number of airports per bracket of number of annual IFR movements.

```{r}
#| label: tbl-apt-classification
#| tbl-cap: Airport classification - 2019

tribble(
  ~movements, ~category, ~airports,
  ">250,000",             "Very large",   "14",
  "[250,000 - 150,000]",  "Large",        "18",
  "[149,999 - 40,000]",   "Medium",       "76",
  "[39,999 - 15,000]",    "Small",        "92",
  "<15,000",              "Other",        "957"
) %>% 
  gt() %>%
  cols_label(
    movements = "Annual IFR movements",
    category = "Airport category",
    airports = "Number of airports"
    ) %>% 
  cols_align(
    align = "right",
    columns = c(airports)
  ) %>% 
  cols_align(
    align = "left",
    columns = c(movements:category)
  ) %>% 
  tab_footnote(
    footnote = "A movement is either a take off or a landing at an airport",
    locations = cells_column_labels(
      columns = movements
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: Data compiled by SESAR 2020 experts on the basis of data provided by the EUROCONTROL Performance Review Unit (PRU), ECAC States*")
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(33),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold",
    column_labels.background.color = "#003366"
  )
```

## Description

An airport can be classified in several ways. Here we focus on categorisation of
airports according to the number of IFR movements.

The list of airports used for this classification was developed using a two-step 
procedure: 

**Step 1:** The initial worldwide airport list provided by the EUROCONTROL 
Performance Review Unit (PRU) was restricted to airports located in ECAC Member 
States and having both ICAO and IATA codes, in order to focus on airports providing 
commercial air transport services (1 079 airports). 

**Step 2:** Seventy-seven airports located in ECAC Member States were added to
the airport list in order to scope all airports for which the EUROCONTROL PRU
provided operational data.

:::{.callout-note}
Statistics on individual airport movements and operations at Airports can be 
downloaded from the [Aviation Intelligence Unit Dashboard](https://ansperformance.eu/data/).
:::

## Other possible sources

Information on airports that have implemented a collaborative decision-making (A-CDM) 
can be found [via this link](https://www.eurocontrol.int/concept/airport-collaborative-decision-making). 

Airport CDM (A-CDM) aims to improve the overall efficiency of airport operations
by optimising the use of resources and improving the predictability of events. 
It focuses especially on aircraft turnaround and pre-departure sequencing processes.

The A-CDM concept has been globally recognised. A-CDM is fully implemented in 30 
airports across Europe (status: December 2020), including Amsterdam, Barcelona,
Bergamo, Berlin Brandenburg "Willy Brandt", Brussels, Copenhagen, Düsseldorf, 
Frankfurt, Geneva, Hamburg, Helsinki, Lisbon, London Gatwick, London Heathrow, 
Lyon, Madrid, Milan Malpensa, Milan Linate, Munich, Naples, Nice, Paris CDG, 
Paris Orly, Oslo, Palma de Mallorca, Prague, Rome Fiumicino, Stockholm Arlanda, 
Stuttgart, Venice, Warsaw and Zurich.

More details for a selected airport are available in the [EUROCONTROL Public Airport Corner](https://www.eurocontrol.int/tool/airport-corner)

## Comment

The mapping of airports to categories in the recommended value is purely indicative
and is based on the situation in 2018. The mapping of airports will most probably 
change significantly as a result of COVID from 2020 onwards. The local situation
of many airports may not be known or be interpreted differently. Final applicability 
of the assigning of airports to categories needs to be checked and confirmed by 
the appropriate airport or authority.

## Related inputs

@sec-turnaround-time [Turnaround time](#sec-turnaround-time)



