# Amount of emissions released by fuel burn {#sec-amount-of-emissions-released-by-fuel-burn}


```{r}
library(readxl)
library(dplyr)
library(gt)
library(tidyverse)
```

## EUROCONTROL recommended values

This input represents the amount of emissions produced by combustion of aviation
fuel, focusing on the main types of pollutants.

```{r}
#| label: tbl-emissions-per-kg-fuel
#| tbl-cap: Estimated amount of emissions released per kg of fuel burnt

tribble(
  ~type, ~emissions,
  "CO2"        , "3.15 kg Kerosene",
  "H2O"        , "1.237 kg",
  "SOx"        , "0.00084 kg",
  ""           , "0.000114 kg",
  "NOx"        , "0.0148 kg",
  "HC"         , "0.00032 kg",
  "CO"         , "0.00325 kg",
  "PM volatile", "0.000092 kg"
) %>% 
  gt() %>% 
  tab_header(title = "Emissions per kg of fuel burnt") %>%
  cols_align(
    align = "right",
    columns = "emissions"
  ) %>% 
  cols_label(
    type = md("**Emission type**"),
    emissions = md("**Emission**")
  ) %>%
  # tab_style(
  #   style = list(
  #     cell_text(weight = "bold")
  #   ),
  #   locations = cells_body(
  #     columns = "type"
  #   )
  # ) %>% 
  # tab_spanner(
  #   "Emissions per kg of fuel burnt",
  #   1:2
  # ) %>% 
  tab_footnote(
    footnote = "Source (1)",
    locations = cells_body(columns = emissions, rows = 1:2),
    placement = "right"
  ) %>% 
  tab_footnote(
    footnote = "assuming 440 ppm, Source (2)",
    locations = cells_body(columns = emissions, rows = 3),
      placement = "right"
  ) %>% 
  tab_footnote(
    footnote = "assuming 600 ppm, Source (2)",
    locations = cells_body(columns = emissions, rows = 4),
      placement = "right"
  ) %>% 
  tab_footnote(
    footnote = "Source (3)",
    locations = cells_body(columns = emissions, rows = 5:8),
    placement = "right"
  ) %>% 
  tab_options(
    table.width = pct(70),
    footnotes.marks = c(letters[1:10])
   )
```
*Sources:*

*(1) - EUROCONTROL, "European Aviation Fuel Burn and Emissions Inventory System for the European Environment Agency ", 2018[@EuropeanAviationFuel2018]*

*(2) - MIT, "Laboratory for aviation and the environment – Guidance on the use of AEDT for gridded aircraft emissions in atmospheric models", 2010[@LaboratoryAviationEnvironment2010]*

*(3) - EUROCONTROL, EASA, EU, "European Aviation Environmental Report, EU27+FTA aviation emissions for year 2021", 2022[@eaer2022]*

## When to use the input? 

This input is recommended for a wide use in assessments that focus on the 
assessment of environmental impact from the burning of fuel at any stage of the 
flight.

## Comments

The Committee on Aviation Environmental Protection (CAEP), a technical committee 
of the ICAO Council, recommends the use a conversion factor of **3.16 g of CO~2~ per gram of Jet A.** 
The 3.16 value can be found in ICAO Doc 9889, 1st edition, 2011, 
and other documents. 

However, in Europe, as early as 2009, Commission Decision 2009/339/EC indicated 
an emission **factor of 3.15 for the mass conversion from Jet A to CO~2~** for the
period after January 2021.

In view of the above, emission factor 3.15 should continue to be used in SESAR 
2020, for the sake of internal consistency within the programme, unless the EU 
ETS decides to move to 3.16. 
Factor 3.16 should be used when the evaluation concerns comparisons with studies
carried out within the ICAO framework or using the factor recommended by ICAO, 
in order to ensure external consistency.[@SESAR2020Environment]

## Related inputs

* @sec-rate-of-fuel-burn [Rate of fuel burn](#sec-rate-of-fuel-burn)

* @sec-cost-of-emissions [Cost of emissions](#sec-cost-of-emissions)

* @sec-ifr-flight-information-per-market-segment [IFR flight information per market segment](#sec-ifr-flight-information-per-market-segment)

## Sources interesting to consult

[European Aviation Environmental Report Series](https://www.eurocontrol.int/publication/european-aviation-environmental-report-2022). 
The latest 2022 edition shows the aviation sector has taken steps to address 
environmental challenges but also that more decisive actions are needed. In 
particular the latest trends in aircraft engine emissions can be checked in 
section 3.2. of the document.

[European Environment Agency “EMEP/EEA air pollutant emission inventory guidebook – 2019”](https://www.eea.europa.eu/publications/emep-eea-guidebook-2019) 
provides values for emission factors and fuel consumption in different phases of 
flight – including taxiing – for different aircraft types, using three different
levels of accuracy and complexity, in section 1.A.3.a of the document.

**ICAO Aircraft Engine Emissions Databank**, available at [EASA website](https://www.easa.europa.eu/en/domains/environment/icao-aircraft-engine-emissions-databank),
contains information on exhaust emissions of production aircraft engines, covering 
engine types which emissions are regulated, namely turbojet and turbofan engines
with a static thrust greater than 26.7 kilonewtons.

[Swiss Federal Office of Civil Aviation, FOCA – Aircraft Engine Emissions](https://www.bazl.admin.ch/bazl/en/home/specialists/regulations-and-guidelines/environment/pollutant-emissions/aircraft-engine-emissions.html) 
presents a measurement and calculation methodology for aircraft piston engine
emissions in order to improve aviation emission inventories, as developed by FOCA.

[Swedish Defence Research Agency](https://www.foi.se/en/our-knowledge/aeronautics-and-air-combat-simulation/fois-confidential-database-for-turboprop-engine-emissions.html) holds a database of emission indices of NOx, HC and CO, 
with corresponding fuel flows for turboprop engines.

## References
