# Amount of emissions released by fuel burn {#sec-amount-of-emissions-released-by-fuel-burn}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

This input represents the amount of emissions produced by combustion of aviation
fuel, focusing on the main types of pollutants.

```{r}
#| label: tbl-emissions-per-kg-fuel
#| tbl-cap: Emissions per kg of fuel burnt

read_xlsx("data/emissions_from_fuel.xlsx") %>% 
gt() %>% 
  cols_label(
    type = "Pollutant",
    emissions = "Amount of emissions (kg)"
  ) %>% 
  cols_align(
    align = "center",
    columns = "emissions"
  )
```

*Source: [EUROCONTROL (2018). European Aviation Fuel Burn and Emissions Inventory System for the European Environment Agency](https://www.eurocontrol.int/publication/european-aviation-fuel-burn-and-emissions-inventory-system-feis-european-environment)*

## When to use the input?

This input is recommended for a wide use in assessments that focus on the 
assessment of environmental impact from the burning of fuel at any stage of the 
flight.

## Comments

The Committee on Aviation Environmental Protection (CAEP), a technical committee
of the ICAO Council, recommends the use a conversion factor of **3.16 g of CO2 per
gram of Jet A.** The 3.16 value can be found in ICAO Doc 9889, 1st edition, 2011,
and other documents. 

However, in Europe, as early as 2009, Commission Decision 2009/339/EC indicated
an **emission factor of 3.15 for the mass conversion from Jet A to CO~2~** for the
period after January 2021.

In view of the above, emission factor 3.15 should continue to be used in SESAR 
2020, for the sake of internal consistency within the program, unless the EU 
ETS decides to move to 3.16. 

Factor 3.16 should be used when the evaluation concerns comparisons with studies
carried out within the ICAO framework or using the factor recommended by ICAO, 
in order to ensure external consistency.

## Related inputs

* @sec-rate-of-fuel-burn [Rate of fuel burn](#sec-rate-of-fuel-burn)

* @sec-cost-of-emissions [Cost of emissions](#sec-cost-of-emissions)

* @sec-ifr-flight-information-per-market-segment [IFR flight information per market segment](#sec-ifr-flight-information-per-market-segment)

## References

[@eaer2022][@eea:2019][@easaICAOAircraftEngine]
