# ANSP Employment Costs {#sec-ansp-employment-cost}

```{r}
library(readxl)
library(dplyr)
library(gt)
```


## EUROCONTROL recommended values

In this section are considered ANSPs’ average annual employment costs for one Full Time Equivalent (FTE) in euros by category of staff. The table below shows the recommended values in 2020.

```{r}
#| label: tbl-ansp-employment-cost-general
#| tbl-cap: Average yearly ANSPs employment costs

read_xlsx("data/ansp_employment.xlsx",
          sheet = "general") %>%
  gt()%>%
  cols_label(
    funct = "Staff function",
    ectrl_area = "EUROCONTROL area",
    ses_area = "SES area"
  )%>%
  fmt_currency(
    currency = "EUR",
    columns = ectrl_area:ses_area,
    decimals = 0
  )%>%
   tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = funct:ses_area,
      rows = 3
    )
   )
```
*Source: EUROCONTROL, "ATM cost-effectiveness (ACE) benchmarking report for 2020 with 2021-2024 outlook", 2022 [@acereports]*


## Description

One full-time equivalent (FTE) is a single person carrying out a particular job 
or activity working on a full-time basis during a year. A part-time employee
working half-time would be counted as 0.5 FTEs. A full-time ATCO working two 
thirds of their time on duty in OPS and one third of their time on teaching at 
a training academy would be counted as 0.67 FTE under ‘ATCOs in OPS’ and a 0.33 
FTE under other categories as relevant.

Employment costs comprise gross wages and salaries, payment for overtime, employer 
contributions to social security schemes and taxes, pension contributions and
other benefits. For a study on employment costs, the categories of staff working
in an ANSP have been divided into two:


* **ATCOs in ops:** ATCOs participating in an activity which is either directly 
related to the control of traffic or where there is a necessary requirement for 
ATCOs to be able to control traffic

* **Support staff or non-ATCO in ops:** this category includes all other staff.
It includes ATCOs on other duties (participating in an activity outside OPS, such
as special projects, teaching at a training academy, providing instruction in a 
simulator, working in a full-time management position, etc.), trainees, ATC 
assistants, technical and operational support staff, administration staff, and others.

The following table gives an overview of individual ANSPs and average European 
system FTE costs for the two categories. The values were calculated based on 
information extracted from EUROCONTROL ACE Report[@ace2020].

```{r}
#| label: tbl-ansp-employment-cost-detailed
#| tbl-cap: Individual ANSP and average European system FTE costs

read_xlsx("data/ansp_employment.xlsx",
          sheet = "all") %>%
  gt()%>%
  cols_label(
    ansp = "ANSP",
    atco_ops = "ATCO in ops",
    support = "Support staff",
    all = "All staff"
  )%>%
  fmt_currency(
    currency = "EUR",
    columns = atco_ops:all,
    decimals = 0
  )%>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = ansp:all,
      rows = 39
    )
  )
```
*Source: EUROCONTROL, "ATM cost-effectiveness (ACE) benchmarking report for 2020 with 2021-2024 outlook", 2022 [@acereports]*

The values in the table above were calculated using values provided in Annex 5 – 
Tables 0.3 and 0.5 of the source document[@acereports].

:::{.callout-note}
The employment costs above refer to gate-to-gate cost (i.e. en route and terminal 
costs) and are expressed in 2020 prices.
:::

## Related inputs
* @sec-atm-cost-effectiveness-indicators [ATM cost-effectiveness indicators](#sec-atm-cost-effectiveness-indicators)

## References

