# ANSP Employment Costs {#sec-ansp-employment-cost}

```{r}
library(readxl)
library(dplyr)
library(gt)
```


## EUROCONTROL recommended values

In this section are considered ANSPs’ average annual employment costs for one Full Time Equivalent (FTE) in euros by category of staff. The table below shows the recommended values in 2020.

```{r}
#| label: tbl-ansp-employment-cost-general
#| tbl-cap: Average yearly ANSPs employment costs

read_xlsx("data/ansp_employment.xlsx",
          sheet = "general") %>%
  gt()%>%
  cols_label(
    funct = "Staff function",
    ectrl_area = "EUROCONTROL area",
    ses_area = "SES area"
  )%>%
  fmt_currency(
    currency = "EUR",
    columns = ectrl_area:ses_area,
    decimals = 0
  )%>%
   tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = funct:ses_area,
      rows = 3
    )
   )%>%
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    funct ~ px(50),
    ectrl_area ~ px(25),
    ses_area ~ px(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source = md("*Source: [EUROCONTROL (2022), ATM cost-effectiveness (ACE) benchmarking report for 2020 with 2021-2024 outlook](https://ansperformance.eu/publications/prc/ace/)*")
  )
```

## Description

One full-time equivalent (FTE) is the equivalent of a single person carrying out
a particular job or activity working on a full-time basis during a year. 
A part-time employee working half-time would be counted as 0.5 FTEs. A full-time 
ATCO working two thirds of her time on duty in ops and one third of her time on
teaching at a training academy would be counted as a 0.67 FTE ATCO in ops and a 
0.33 FTE ATCO on other duties.

Employment costs comprise gross wages and salaries, payment for overtime, employer
contributions to social security schemes and taxes, pension contributions and other 
benefits. For a study on employment costs, the categories of staff working in an 
ANSP have been divided into two:

* **ATCOs in ops:** ATCOs participating in an activity which is either directly 
related to the control of traffic or where there is a necessary requirement for 
ATCOs to be able to control traffic

* **Support staff or non-ATCO in ops:** this category includes all other staff. 
It includes ATCOs on other duties (participating in an activity outside ops, 
such as special projects, teaching at a training academy, providing instruction 
in a simulator, working in a full-time management position, etc.), trainees, ATC
assistants, technical and operational support staff, administration staff, and others

The following table gives an overview of individual ANSPs and average European 
system FTE costs for the two categories. The values were calculated based on 
information extracted from EUROCONTROL ACE Report.[@ace2020]

```{r}
#| label: tbl-ansp-empl-cost
#| tbl-cap: Individual ANSP and average European system FTE costs

read_xlsx("data/ansp_employment.xlsx",
          sheet = "all") %>%
  gt()%>%
  cols_label(
    ansp = "ANSP",
    atco_ops = "ATCO in ops",
    support = "Support staff",
    all = "All staff"
  )%>%
  fmt_currency(
    currency = "EUR",
    columns = atco_ops:all,
    decimals = 0
  )%>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = ansp:all,
      rows = 39
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source = md("*Source: [EUROCONTROL PRU (2022), Air traffic management cost-effectiveness (ACE) benchmarking report for 2020 with 2021-2024 outlook](https://ansperformance.eu/publications/prc/ace/)*")
  )
```

The values in @tbl-ansp-empl-cost were calculated using values provided in 
Annex 5, Tables 0.3 and 0.5 of the source document.[@ace2020]

:::{.callout-note}
The employment costs above refer to gate-to-gate cost (i.e. en route and terminal 
costs) and are expressed in 2020 prices.
:::

## Related inputs
* @sec-atm-cost-effectiveness-indicators [ATM cost-effectiveness indicators](#sec-atm-cost-effectiveness-indicators)

## References

