# ATM operational units {#sec-atm-operational-units}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

This value represents the number of ATC units (air traffic centers) providing ATC services across Europe for the purpose of:

* preventing collisions between aircraft and on the maneuvering area between aircraft and obstructions

* expediting and maintaining an orderly flow of air traffic.

```{r}
#| label: tbl-atm-op-units
#| tbl-cap: Number of ATM operational units in Europe

read_xlsx("data/atm_operational_units.xlsx")%>% 
  gt() %>% 
  cols_label(
    year = "Year",
    ansp = "ANSP",
    acc = "ACC",
    app = "APP",
    twr = "TWR",
    afis = "AFIS units",
    atc = "ATC sectors"
  ) %>% 
  cols_align(
    align = "center",
    columns = c(ansp:atc)
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold"),
      cell_fill(color = "#FFFAFA")
    ),
    locations = cells_body(
      columns = "year"
    )
  ) %>% 
  tab_footnote(
    footnote = "ANSP - Air Navigation Service Provider",
    locations = cells_column_labels(
      column = ansp
    )
  ) %>% 
  tab_footnote(
    footnote = "ACC - Area Control Centre",
    locations = cells_column_labels(
      column = acc
    )
  ) %>% 
  tab_footnote(
    footnote = "APP - Approach Units",
    locations = cells_column_labels(
      column = app
    )
  ) %>% 
  tab_footnote(
    footnote = "TWR - Tower",
    locations = cells_column_labels(
      column = twr
    )
  ) %>% 
  tab_footnote(
    footnote = "AFIS - Airport/Aerodrome Flight Information Service",
    locations = cells_column_labels(
      column = afis
    )
  ) %>% 
  tab_footnote(
    footnote = "ATC - Air Traffic Control. Represents the number of sectors open at maximum configuration",
    locations = cells_column_labels(
      column = atc
    )
  )
    ```
*Source: EUROCONTROL, “ATM Cost-Effectiveness (ACE) Benchmarking Reports – Annex 5” [@acereports]*

## Description

The air traffic control airspace management and air traffic flow management services which make up ATM are the main services provided by national ANSPs. Generally there is one service provider per State, except in a few cases where one ANSP provides ATC services in several States.^[Belgocontrol for Belgium and Luxembourg, SMATSA for Serbia and Montenegro, the DSNA for France and Monaco, and
MUAC for the upper airspace of Benelux and northern Germany]

The ATM Cost Effectiveness (ACE) Benchmarking Report presents data about and analysis of cost effectiveness and productivity for ANSPs. It presents a review and comparison of ATM cost effectiveness for the 38 air navigation service providers (ANSPs) in Europe, which provide coverage for EUROCONTROL’s 41 Member States and 2 Comprehensive Agreement States.^[Morocco and Israel each signed a Comprehensive Agreement in 2016]

It excludes, however, elements related to services provided to military operational air traffic (OAT), oceanic ANS, and landside airport management operations.


## When to use this input?

The values presented above can be used in analyses where the level of granularity of information per type of provider or per provider is important. This is typically interesting for studies looking into costs or benefits per ANSP/ACC/Tower/other.

## References
