# CNS infrastructure {#sec-cns-infrastructure}

```{r}
library(readxl)
library(dplyr)
library(gt)
```


## EUROCONTROL recommended values

The values below show the number of civil owned systems installed which are 
devoted to carrying out navigation and surveillance functions in ECAC.

### Navigation aids

@tbl-cns-ecac1 and @tbl-cns-ecac2 show the number of Navigation aids in ECAC as per end of 2022. 
Please note there might be some marginal differences with the latest available
picture in the [EUROCONTROL GNI tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome) depending on the date of extraction.

```{r}
#| label: tbl-cns-ecac1
#| tbl-cap: Navigation aids in the ECAC Member States - part 1

read_xlsx("data/cns_infrastructure.xlsx",
          sheet = "navaids1") %>% 
  gt() %>% 
  cols_align(
    align = "right",
    columns = everything()
  ) %>% 
  cols_align(
    align = "left",
    columns = "State"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "State"
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(10)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 45
    )
  ) %>% 
  tab_footnote(
    footnote = "Considered in Serbia as service is provided by SMATSA",
    locations = cells_body(
      columns = State,
      rows = 28
    )
  ) %>% 
   tab_footnote(
    footnote = "Including Azores and Madeira",
    locations = cells_body(
      columns = State,
      rows = 33
    )
  ) %>% 
  tab_footnote(
    footnote = "Considering jointly the State of Serbia and the State of Montenegro",
    locations = cells_body(
      columns = State,
      rows = 36
    )
  ) %>% 
   tab_footnote(
    footnote = "Including Canary Islands; Ceuta and Melilla and Spain Continental",
    locations = cells_body(
      columns = State,
      rows = 39
    )
  ) %>% 
  tab_source_note (
    source = md("*[EUROCONTROL Ground-based Navigation Infrastructure Map Tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)*")
  )
```

```{r}
#| label: tbl-cns-ecac2
#| tbl-cap: Navigation aids in the ECAC Member States - part 2

read_xlsx("data/cns_infrastructure.xlsx",
          sheet = "navaids2") %>% 
  gt() %>% 
  cols_align(
    align = "right",
    columns = everything()
  ) %>% 
  cols_align(
    align = "left",
    columns = "State"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "State"
    )
  ) %>% 
  tab_spanner(
    "ILS",
    columns = 2:4
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(20)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 45
    )
  ) %>% 
  tab_footnote(
    footnote = "Considered in Serbia as service is provided by SMATSA",
    locations = cells_body(
      columns = State,
      rows = 28
    )
  ) %>% 
   tab_footnote(
    footnote = "Including Azores and Madeira",
    locations = cells_body(
      columns = State,
      rows = 33
    )
  ) %>% 
  tab_footnote(
    footnote = "Considering jointly the State of Serbia and the State of Montenegro",
    locations = cells_body(
      columns = State,
      rows = 36
    )
  ) %>% 
  tab_footnote(
    footnote = "Including Canary Islands; Ceuta and Melilla and Spain Continental",
    locations = cells_body(
      columns = State,
      rows = 39
    )
  ) %>% 
  tab_source_note (
    source = md("*[EUROCONTROL Ground-based Navigation Infrastructure Map Tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)*")
  )
```

The [EUROCONTROL Ground-based Navigation Infrastructure Map Tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)
illustrates the current status and the evolution plans for the Ground-based 
Navigation Infrastructure in ECAC. The navaids database is updated based on the 
information published in the [European AIS Database (EAD)](https://www.eurocontrol.int/service/european-ais-database) and the planning 
provided by Air Navigation Service Providers. Please note the evolution plans 
shall not be considered as a commitment by a State / ANSP.

### Surveillance units

The Table below shows the number of Surveillance units in ECAC as per September 2020. 
Please note work is in progress to get more recent figures.

```{r}
#| label: tbl-cns-ecac-surv
#| tbl-cap: Surveillance aids in EUROCONTROL Member States

read_xlsx("data/cns_infrastructure.xlsx",
          sheet = "surveillance") %>% 
  gt() %>% 
  cols_align(
    align = "right",
    columns = everything()
  ) %>% 
  cols_align(
    align = "left",
    columns = "State"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "State"
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    State ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 45
    )
  ) %>% 
  tab_footnote(
    footnote = "Considered in Serbia as service is provided by SMATSA",
    locations = cells_body(
      columns = State,
      rows = 28
    )
  ) %>% 
   tab_footnote(
    footnote = "Including Azores and Madeira",
    locations = cells_body(
      columns = State,
      rows = 33
    )
  ) %>% 
  tab_footnote(
    footnote = "Considering jointly the State of Serbia and the State of Montenegro",
    locations = cells_body(
      columns = State,
      rows = 36
    )
  ) %>% 
  tab_footnote(
    footnote = "Including Canary Islands; Ceuta and Melilla and Spain Continental",
    locations = cells_body(
      columns = State,
      rows = 39
    )
  ) %>% 
  tab_source_note (
    source = md("*Source: [EUROCONTROL Ground-based Navigation Infrastructure Map Tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)*")
  )
```

The Mode S, PSR (primary surveillance radar) and SSR (secondary surveillance radar)
numbers are extracted from the surveillance database of the EUROCONTROL CNS unit.

The data depicted in @tbl-cns-ecac-surv above shows the status as per September 2020. As
the allocation and implementation of Mode S interrogator codes (ICs) require a 
coordinated approach, every installation of a Mode S radar is officially registered.
The numbers of PSRs and SSRs reported above are not necessarily accurate, as 
they are based on voluntary reports by the Member States on updates and changes 
to their surveillance infrastructure. Work on the collection of MLAT/ADS-B 
stations is still in progress.

The WAM/ADS B (wide-area multilateration/automatic dependent surveillance-broadcast)
and ADS B data originate from the database, which is maintained by the EUROCONTROL
Surveillance and Code Coordination Unit and is based on inputs from stakeholders.

The ADS B and WAM Section coordinates the deployment of initial ADS B applications 
and WAM in Europe. The WAM/ADS-B sensor count only includes sensors mainly used 
for surveillance of airborne aircraft (e.g. in TMAs or en route). It does not 
include sensors mainly used for airport surface surveillance (e.g. airport MLAT 
used for A-SMGCS). 

:::{.callout-note}
Please note that the above data were provided by States and may slightly differ 
from the values in the table “PBN and precision approach procedures” as they are
valid for different reference periods.
:::

## Related inputs

@sec-fleet-cns-capability [Fleet CNS capability](#sec-fleet-cns-capability)

@sec-pbn-and-precision-approach-procedures [PBN and precision approach procedures](#sec-pbn-and-precision-approach-procedures)
