---
editor_options: 
  markdown: 
    wrap: 72
---

# General parameters {#sec-conversions .unnumbered}

Below are presented some key values used in the Standard Inputs and that can serve as a reference for other uses.

## Inflation {#inflation .unnumbered}

The inflation levels have been calculated based on the Harmonised Index
of Consumer Prices (HICP), as the yearly difference in the HICP. The
HICP data was extracted from [Eurostat](http://ec.europa.eu/eurostat)
HICP - annual data (average index and rate of change) database
(prc_hicp_aind).The annual change in the index is shown in @tbl-inflation-table.

```{r}
#| label: tbl-inflation-table
#| tbl-cap: Annual average inflation values

library(readxl)
library(dplyr, quietly = TRUE)
library(gt)

# read from CSV (exported directly from Eurostat, see R/export_inflation.R)
readxl::read_xlsx("data/values_2022.xlsx",
                  sheet = "basic assumptions",
                  range = "C6:E33" 
                  ) %>%
  filter(2000 <= Year, Year <= 2021) %>%
  arrange(desc(Year)) %>% 
  # format the table
  gt() %>%
  tab_source_note(
    source_note = md("*Source: [Eurostat](http://ec.europa.eu/eurostat)*")
  ) %>%
  cols_label(
    Year  = "Year",
    Index = "Index",
    Rate  = "Rate of change"
  ) %>%
  fmt_percent(
    columns = c(Rate),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(Index),
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = c(Year, Index, Rate)
  )
```

## Exchange Rate Conversion {#exchange-rates .unnumbered}

The currency exchange rates provided below are based on the European Central Bank (ECB) rates that are published daily. The rates shown in the table below correspond to the average rate for the year 2021 as published by ECB.

```{r }
#| label: tbl-eurostat-exchange-rates
#| tbl-cap: Average euro foreign exchange rate (2021)
library(readr)
library(dplyr)
library(gt)

read_csv("data/exchange_rate.csv", show_col_types = FALSE) %>%
  gt() %>%
  cols_label(
    currency  = "Currency",
    to_eur    = "Currency-€",
    from_eur  = "€-Currency"
  ) %>%
    tab_source_note(
    source_note = md("*Source: [European Central Bank](https://www.ecb.europa.eu/stats/policy_and_exchange_rates/euro_reference_exchange_rates/html/index.en.html)*")
  ) %>%
  cols_align(
    align = "center",
    columns = c(to_eur, from_eur)
  )
```

## Cost of fuel {.unnumbered}

The cost of fuel used in this document is based on the 2021 average jet
fuel price handled by [IATA](https://www.iata.org/en/publications/economics/fuel-monitor/), unless otherwise specified. All conversions are done using the values specified on this page.

```{r}
#| label: tbl-cost-of-fuel
#| tbl-cap: Average jet fuel price
library(dplyr)
library(gt)

tribble(
  ~currency, ~price_barrel, ~price_gallon, ~price_kg,
  "USD", 141.7, 3.8, 1.1,
  "EUR", 138.9, 3.3, 1.1,
) %>%
  gt() %>%
  cols_label(
    currency = "Currency",
    price_barrel = "Price per barrel",
    price_gallon  = "Price per gallon",
    price_kg = "Price per kg"
  ) %>%
    tab_source_note(
    source_note = md("*Source: [IATA](https://www.iata.org/en/publications/economics/fuel-monitor/)*")
  ) %>%
  cols_align(
    align = "center",
    columns = c(price_barrel, price_gallon, price_kg)
  )
```

## Conversion values {.unnumbered}

```{r}
#| label: tbl-conversion-values
#| tbl-cap: Units conversion values
tribble(
  ~from, ~to,
  "1 nautical mile (NM)", "1.852 km",
  "1 kilometre (km)"    , "0.53996 NM",
  "1 tonne (metric = 1 000 kg) of jet fuel", "325.33 US gallons",
  ""                                       , "1 235 litres",
  ""                                       , "7.8 barrels",
  "1 barrel (bbl) of jet fuel", "42 US gallons",
  ""                          , "158.99 litres",
  ""                          , "0.1291 tonne = 129.10 kg",
  "1 US gallon of jet fuel (US gal)", "3.7854 litres",
  ""                                , "3.073 kg",
  "", "6.7764 lb",
  "Density of kerosene", "0.812 kg/litre",
  "1 litre of fuel (l)", "0.26417 US gallons",
  "1 kilogramme of fuel (kg)", "2.2046 lb",
  "1 pound of fuel (lb)", "0.45359 kg"
) %>% 
  gt() %>%
  cols_label(
    from = "From",
    to = "To",
    )
```
