# Conversions, Inflation, Cost of Fuel, Exchange Rate {.unnumbered #conversions}


## Inflation {.unnumbered #inflation}

All values are given in euro (€) at December 2016 price levels (unless otherwise indicated). They
have been inflated from values given in the source documents by means of the
[EUROSTAT European Union][eu-enlargement]
(EC6-1972, EC9-1980, EC10-1985, EC12-1994, EU15-2004, EU25-2006,
EU27-2013, EU28) harmonised index of consumer prices (HICP 2015=100).

The annual change in
the index is shown below. The values of the index are available from [Eurostat][eurostat]
data set in the [`prc_hicp_aind` table][prc_hicp_aind].

[eurostat]: <http://ec.europa.eu/eurostat> 
[prc_hicp_aind]: <https://ec.europa.eu/eurostat/data/database?node_code=prc_hicp_aind>
[eu-enlargement]: <https://ec.europa.eu/eurostat/statistics-explained/index.php/Glossary:EU_enlargements>

(ref:inflation-table) Annual average inflation values.


```{r inflation-table, , fig.cap='(ref:inflation-table)'}
library(readr)
library(dplyr)
library(gt)

# read from CSV (exported directly from Eurostat, see R/export_inflation.R)
read_csv("data/inflation.csv") %>%
  # format the table
  gt() %>%
  tab_source_note(
    source_note = md("Source: Eurostat, [http://ec.europa.eu/eurostat](http://ec.europa.eu/eurostat)")
  ) %>%
  cols_label(
    year  = "Year",
    index = "Index",
    rate  = "Rate of change"
  ) %>%
  fmt_percent(
    columns = vars(rate),
    decimals = 2
  ) %>%
  fmt_number(
    columns = vars(index),
    decimals = 2
  )
```


## Exchange Rate Conversion {.unnumbered #exchange-rates}

Values in pounds sterling (GB £) and US dollars (US $) have been converted to euros (€) using the
2016 average euro foreign exchange rate.
Currency Currency: € €: Currency

```{r eurostat-exchange-rates}
library(readr)
library(dplyr)
library(gt)

read_csv("data/exchange_rate.csv") %>%
  gt() %>%
  cols_label(
    currency  = "Currency",
    to_eur    = "Currency: €",
    from_eur  = "€: Currency"
  )
```


## Cost of Fuel {.unnumbered #cost-of-fuel}


## Conversion Values {.unnumbered #conversion-values}


## Further Information {.unnumbered}




