# Conversions, Inflation, Cost of Fuel, Exchange Rate {.unnumbered #conversions}


## Inflation {.unnumbered #inflation}

All values are given in euro (€) at December 2019 price levels (unless otherwise indicated).
They have been inflated from values given in the source documents by means of the
[EUROSTAT European Union][eu-enlargement]
(EU6-1958, EU9-1973, EU10-1981, EU12-1986, EU15-1995, EU25-2004, EU27- 2007,
EU-28-2013, EU27-2020) harmonised index of consumer prices (HICP 2015=100).

The annual change in the index is shown below.
The values of the index are available from [Eurostat][eurostat]
(table [`prc_hicp_aind` table][prc_hicp_aind].)

[eurostat]: <http://ec.europa.eu/eurostat> 
[prc_hicp_aind]: <https://ec.europa.eu/eurostat/data/database?node_code=prc_hicp_aind>
[eu-enlargement]: <https://ec.europa.eu/eurostat/statistics-explained/index.php/Glossary:EU_enlargements>

(ref:inflation-table) Annual average inflation values.


```{r inflation-table, fig.cap='(ref:inflation-table)'}
library(readxl)
library(dplyr)
library(gt)

# read from CSV (exported directly from Eurostat, see R/export_inflation.R)
readxl::read_xlsx("data/Values 2019_Review.xlsx",
                  sheet = "basic assumptions",
                  range = "C6:E33" 
                  ) %>%
  filter(2000 <= Year, Year <= 2019) %>%
  arrange(desc(Year)) %>% 
  # format the table
  gt() %>%
  tab_source_note(
    source_note = md("Source: Eurostat, [http://ec.europa.eu/eurostat](http://ec.europa.eu/eurostat)")
  ) %>%
  cols_label(
    Year  = "Year",
    Index = "Index",
    Rate  = "Rate of change"
  ) %>%
  fmt_percent(
    columns = vars(Rate),
    decimals = 2
  ) %>%
  fmt_number(
    columns = vars(Index),
    decimals = 2
  )
```


## Exchange Rate Conversion {.unnumbered #exchange-rates}

Values in pounds sterling (GB £) and US dollars (US $) have been converted
to euros (€) using the 2019 average euro foreign exchange rate.

```{r eurostat-exchange-rates}
library(readr)
library(dplyr)
library(gt)

read_csv("data/exchange_rate.csv") %>%
  gt() %>%
  cols_label(
    currency  = "Currency",
    to_eur    = "Currency-€",
    from_eur  = "€-Currency"
  )
```

Further information can be found in the related standard input exchange rate (page 123).


## Cost of fuel

The cost of fuel used in this document is based on the 2019 average jet
fuel price handled by IATA (unless otherwise specified).

```{r cost-of-fuel}
library(dplyr)
library(gt)

tribble(
  ~usd_gallon, ~eur_kg,
  1.83, 0.43
) %>%
  gt() %>%
  cols_label(
    usd_gallon = "US$/gallon",
    eur_kg     = "€/kg",
  )
```

Details can be found in the related standard input
[cost of aviation fuel](#cost_of_aviation_fuel).


## Conversion values

```{r, conversion-values}
tribble(
  ~from, ~to,
  "1 nautical mile (NM)", "1.852 km",
  "1 kilometre (km)"    , "0.53996 NM",
  "1 tonne (metric = 1 000 kg) of jet fuel", "325.33 US gallons",
  ""                                       , "1 235 litres",
  ""                                       , "7.8 barrels",
  "1 barrel (bbl) of jet fuel", "42 US gallons",
  ""                          , "158.99 litres",
  ""                          , "0.1291 ton = 129.10 kg",
  "1 US gallon of jet fuel (US gal)", "3.7854 litres",
  ""                                , "3.073 kg",
  "", "6.7764 lb",
  "Density of kerosene", "0.812 kg/litre",
  "1 litre of fuel (l)", "0.26417 US gallons",
  "1 kilogramme of fuel (kg)", "2.2046 lb",
  "1 pound of fuel (lb)", "0.45359 kg"
) %>% 
  gt() %>%
  cols_label(
    from = "From",
    to     = "To",
  )

```

## For Further Information

For any questions relating to this document, please contact EUROCONTROL using the e-mail [aviation.intelligence@eurocontrol.int](mailto:aviation.intelligence@eurocontrol.int?subject=Standard%20inputs), naming the standard inputs in the subject line.
