# Cost of noise {#sec-cost-of-noise}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

The figures hereafter provide values recommended to be used to estimate the cost 
of noise per person affected, taking into account the cost of annoyance as well
as the health costs due to exposure to air traffic noise.

@tbl-yearly-noise-cost provides an **estimation of the average cost of annoyance for the population, health-related costs, as well as a total (i.e. annoyance and health) from aviation traffic noise for EU27+UK.**

Annoyance refers to the disturbance which individuals experience when they are 
exposed to noise (e.g. discomfort, inconvenience, etc.).

Health impacts are those caused by long-term exposure to noise, such as stress-related 
health problems. Evidence has not been strong for all noise-related health impacts,
and consequently in the European Handbook on External Costs of Transport, only the 
following health impacts are considered: hypertension, ischaemic heart disease,
stroke and dementia. Insomnia is not included in order to avoid double-counting 
with the costs of annoyance.

The data is presented for different noise levels and the values in euros represent 
cost per person and dB.

```{r}
#| label: tbl-yearly-noise-cost
#| tbl-cap: Yearly environmental cost of aviation traffic noise for the EU27+UK
read_xlsx("data/cost_noise.xlsx",
          sheet = "yearly") %>% 
  gt() %>% 
  cols_label(
    noise = "Noise level (Lden in dB(A))",
    annoyance = "Annoyance",
    health = "Health",
    total = "Total"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = annoyance:total,
    decimals = 2
  ) %>% 
  cols_align(
    align = "right",
    columns = annoyance:total
  ) %>% 
  tab_footnote(
    footnote = "The monetary values are adjusted from 2016 to 2022 prices",
    locations = cells_column_labels(
      columns = annoyance:total
    )
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
  cols_width(
    noise ~ pct(25),
    annoyance ~ pct(25),
    health ~ pct(25),
    total ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "Lden is the common EU indicator which corresponds to the average noise level throughout the day, evening and night to which a citizen is exposed over a year. One fundamental feature of Lden is that it assumes that evening and night-time noise is more of a nuisance than daytime noise. 	(Evening noise is given a penalty of 5 dB(A). Night-time noise is given a penalty of 10 dB(A).)",
    locations = cells_column_labels(
      columns = noise
    )
  ) %>% 
  tab_footnote(
    footnote = "The basic measurement index for noise is the decibel (dB). It is indexed logarithmically, reflecting the logarithmic manner in which the human ear responds to sound pressure. Within the human range of hearing, deep and very high tones at the same sound intensity are experienced as less noisy. To correct for this sensitivity, a frequency weighting is applied to measurements and calculations.  The most common frequency weighting is the ‘A weighting’, dB(A).",
    locations = cells_column_labels(
      columns = noise
    )
  )
```
*Source: European Commission, Directorate-General for Mobility and Transport, Essen, H., Fiorello, D., El Beyrouty, K., et al., Handbook on the external costs of transport : version 2019 – 1.1, Publications Office, 2020[@ecdgmove2019]*

@tbl-total-noise-cost presents an assessment of the costs of noise for short, 
medium and long-haul flights based on an analysis of 33 EU airports.

```{r}
#| label: tbl-total-noise-cost
#| tbl-cap: Total and average cost of noise cost for aviation at airports
read_xlsx("data/cost_noise.xlsx",
          sheet = "total") %>% 
  gt() %>%
  cols_label(
    flight = " ",
    total = "Bn €",
    lto = "€ per LTO",
    pax = "€ per pax",
    tonne = "€ per tonne",
    km = "€ per km"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = total:km,
    decimals = 2
  ) %>% 
  cols_align(
    align = "right",
    columns = total:km
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
     flight ~ pct(25),
     total ~ pct(15),
     lto ~ pct(15),
     pax ~ pct(15),
     tonne ~ pct(15),
     km ~ pct(15)
   ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "Costs per pax include the complete flight (not only the half-way principle).",
    locations = cells_column_labels(
      columns = pax
    )
  ) %>% 
  tab_footnote(
    footnote = "The monetary values are adjusted from 2016 to 2022 prices",
    locations = cells_column_labels(
      columns = total:km
    )
  )
```
*Source: European Commission, Directorate-General for Mobility and Transport, Essen, H., Fiorello, D., El Beyrouty, K., et al., Handbook on the external costs of transport : version 2019 – 1.1, Publications Office, 2020[@ecdgmove2019]*

## Other possible values

Below are presented the results of an economic valuation tool developed by the 
UK Department for Environment, Food and Rural Affairs. It converts changes in 
noise exposure to estimated monetary values, in order to support the assessment
of the effects of environmental noise. It details the current understanding of 
the links between environmental noise and various effects, including sleep 
disturbance, annoyance, hypertension and related diseases.[@GuidanceNoisePollution]

```{r}
#| label: tbl-noise-household
#| tbl-cap: Yearly aviation noise marginal cost per household
read_xlsx("data/cost_noise.xlsx",
          sheet = "household") %>% 
  gt() %>%
  cols_label(
    increase = "Increase in daytime noise metric by one decibel (dB)",
    marginal_cost = "Aviation noise marginal cost (excl. sleep disturbance)",
    sleep = "Sleep disturbance cost"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = marginal_cost:sleep,
    decimals = 2
  ) %>% 
  cols_align(
    align = "left",
    columns = increase
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
     everything() ~ pct(33)
   ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "The monetary values are adjusted from 2014 to 2022 prices",
    locations = cells_column_labels(
      columns = marginal_cost:sleep
    )
  )
```
*Source: UK Department for Environment, Food and Rural Affairs, "Environmental Noise: Valuing impacts on: sleep disturbance, annoyance, hypertension, productivity and quiet", 2014[@EnvironmentalNoiseValuing2014]*

## References


