# Cost of emissions {#sec-cost-of-emissions}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

The data provided in the following sub-sections shows an estimation of the cost 
of CO~2~ and other aircraft pollutants released by the combustion of aviation fuel.

### Air pollution

According to the Handbook on the external costs of transport[@ecdgmove2019], for air pollution 
costs, the marginal costs are virtually equal to the average costs. This is due 
to the fact that the dose-response relationships between the emissions of air
pollutants and health effects are nearly linear.

```{r}
#| label: tbl-marginal-pollution-cost
#| tbl-cap: Marginal air pollution costs of aviation
read_xlsx("data/emissions_cost.xlsx",
          sheet = "marginal") %>% 
  gt() %>% 
  cols_label(
    type = "Type of flight",
    distance = "Distance (km)",
    class = "Emissions class",
    aircraft = "Example of aircraft type",
    lto_cost = "Cost per LTO",
    paxkm_cost = "Cost per pax km (€-cent)",
    pax_cost = "Cost per pax"
  ) %>% 
  cols_align(
    align = "right",
    columns = c(lto_cost, paxkm_cost, pax_cost)
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = c(lto_cost:pax_cost),
    decimals = 2
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [European Commission (2019), Handbook on the external costs of transport](https://data.europa.eu/doi/10.2832/51388)*")
  ) %>% 
  tab_footnote(
    footnote = "The monetary values are adjusted to 2022 prices according to inflation",
    locations = cells_column_labels(
      columns = lto_cost:pax_cost
    )
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

### Climate change

One of the approaches to monetise the climate change costs is to estimate the CO~2~ 
cost avoidance, in compliance with the provisions of Paris Climate Agreement. 
@tbl-climate-change-cost provides an estimate of CO~2~ equivalent cost avoidance
for short and medium term. It shows a low, medium and high estimate of these values.

```{r}
#| label: tbl-climate-change-cost
#| tbl-cap: Climate change avoidance costs in € per tonne of CO~2~ equivalent
read_xlsx("data/emissions_cost.xlsx",
          sheet = "climate") %>% 
  gt() %>% 
  cols_label(
    forecast = " ",
    low = "Low",
    medium = "Medium",
    high = "High"
  ) %>% 
  cols_align(
    align = "right",
    columns = low:high
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = low:high
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [European Commission (2019), Handbook on the external costs of transport](https://data.europa.eu/doi/10.2832/51388)*")
  ) %>% 
  tab_footnote(
    footnote = "The values are adjusted to 2022 prices according to inflation",
    locations = cells_column_labels(
      columns = low:high
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
    forecast ~ px(290),
    low ~ px(150),
    medium ~ px(150),
    high ~ px(150)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

### Other possible values

The well-to-tank emissions costs represent the costs linked to the production of
all different type of energy sources, which leads to emissions and other externalities.
It includes the extraction of energy, processing, transport and transmission,
building of energy plants, etc. These emissions are part of the most relevant 
emissions when it comes to transportation.

@tbl-wtt-costs presents the estimated cost of well-to-tank emissions from 
aviation based on an analysis of 33 selected EU airports.

```{r}
#| label: tbl-wtt-costs
#| tbl-cap: Total and average well-to-tank costs for aviation for 33 selected EU airports
read_xlsx("data/emissions_cost.xlsx",
          sheet = "other1") %>% 
  gt() %>% 
  cols_label(
    distance = " ",
    total = "Total cost (bn €)",
    km = "€-cents per pkm",
    pax = "€-cents per pax"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = c(total, km, pax),
    decimals = 1
  ) %>% 
  cols_align(
    align = "right",
    columns = c(total, km, pax)
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [European Commission (2019), Handbook on the external costs of transport](https://data.europa.eu/doi/10.2832/51388)*")
  ) %>% 
  tab_footnote(
    footnote = "The values are adjusted to 2022 prices according to inflation",
    locations = cells_column_labels(
      columns = total:pax
    )
  )  %>% 
  tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
     distance ~ pct(40),
     total ~ pct(20),
     km ~ pct(20),
     pax ~ pct(20)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

@tbl-wtt-pollution presents the damage cost factors used for calculation of the emissions
impacts. The prices are expressed in € per kg and were adjusted to 2022 prices from 2016.

```{r}
#| label: tbl-wtt-pollution
#| tbl-cap: Well-to-tank air pollution costs. Damage cost estimates for EU27+UK
read_xlsx("data/emissions_cost.xlsx",
          sheet = "other2") %>% 
  gt() %>% 
  cols_label(
    cost = " ",
    no = "NOx",
    nmvoc = "NMVOC",
    so = "SO2",
    pm = "PM2.5 (exhaust)"
  ) %>% 
  cols_align(
    align = "right",
    columns = no:pm
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = no:pm,
    decimals = 1
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(140)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [European Commission (2019), Handbook on the external costs of transport](https://data.europa.eu/doi/10.2832/51388)*")
  )
```

## Further reading

Below are listed some sources that may be interesting to consult in the frame of this topic:

* [EUROCONTROL (2022), European Environmental Report](https://www.eurocontrol.int/publication/european-aviation-environmental-report-2022)

* [European Commission Climate Action](https://ec.europa.eu/clima/policies/ets/auctioning_en)

* [UK Department for Environment, Food and Rural Affairs (DEFRA) (2020), Air quality damage cost update 2020](https://uk-air.defra.gov.uk/library/reports?report_id=1006)

## Related inputs

@sec-amount-of-emissions-released-by-fuel-burn [Amount of emissions released by fuel burn](#sec-amount-of-emissions-released-by-fuel-burn)

@sec-rate-of-fuel-burn [Rate of fuel burn](#ec-rate-of-fuel-burn)

## References