# Cost of emissions

## Definition
Estimate of the cost of CO2 and other aircraft emissions released by the
combustion of aviation fuel.

## EUROCONTROL recommended sources and values

Value 1
:   
    ```{r, price-of-co2}
    library(gt)
    library(magrittr)
    library(tibble)
    tribble(
      ~ forecast, ~ low, ~ medium, ~ high,
      "Short and medium run (up to 2030)", 63, 105, 199,
      "Long run (from 2040 to 2060)", 164, 283, 524,
    ) %>% 
      gt() %>% 
      cols_label(
      forecast = "Forecast",
      low = "Low",
      medium = "Medium",
      high = "High"
    ) %>% 
    tab_header(title = md("Price of CO2"),
               subtitle = "Climate change avoidance costs in € per tonne of CO2 equivalent") %>%
    tab_footnote(
      footnote = "adjusted from € 2016 to € 2019 prices",
      locations = cells_column_labels(vars(low, medium, high))) %>% 
    tab_footnote(
      footnote = "These values were derived by calculating the average of
                  the low, central and high estimates for the relevant time
                  periods of the values from the literature, but excluding the
                  lowest and highest values in order to eliminate outliers.",
      locations = cells_column_labels(columns = vars(low))
    ) %>% 
    opt_align_table_header(align = "left")
    ```

Value 2
:   
    
    ```{r, air-pollution-cost}
    library(gt)
    library(magrittr)
    library(tibble)
    tribble(
      ~ cost, ~ nox, ~ nmvoc, ~ so2, ~ pm,
      "EU-27+UK", 11.5, 1.3, 11.5, 20.4
    ) %>% 
      gt() %>% 
      cols_label(
      cost = "Costs in € per kg",
      nox = "NOx",
      nmvoc = "NMVOC",
      so2 = "SO2",
      pm = "High"
    ) %>% 
    tab_header(title = "Well-to-tank air pollution costs: damage cost estimates
              in €/kg emission",
              subtitle = "(emissions in the year 2016, EU-27+UK values)") %>% 
    tab_footnote(
      footnote = "adjusted from € 2016 to € 2019 prices",
      locations = cells_column_labels(vars(nox, nmvoc, so2, pm))) %>% 
      fmt_markdown(columns = TRUE) %>% 
    opt_align_table_header(align = "left")
    ```

Value 3
:   
    ```{r, aviation-air-pollution-cost}
    library(gt)
    library(magrittr)
    library(tibble)
    tribble(
      ~ distance_group, ~ range, ~ eur_year, ~ eur_km, ~ eur_pax,
      "Short-haul",    "< 1 500 km", 284, 0.32, 171,
      "Medium-haul",    "< 1 500-5 000 km", 400, 0.14, 243,
      "Long-haul", "> 5 000 km", 379, 0.06, 467,
    ) %>% 
      gt() %>% 
      cols_label(
      distance_group = "Distance group",
      range = "Range",
      eur_year = "M€/year",
      eur_km = "€-cent/km",
      eur_pax = md("€-cent/ pax<br>(complete flight)")
    ) %>% 
    tab_header(title = md("Total and average air pollution costs for aviation
                          for 33 selected EU airports")) %>%
    tab_footnote(
      footnote = "adjusted from € 2016 to € 2019 prices",
      locations = cells_column_labels(vars(eur_year, eur_km, eur_pax))) %>% 
    tab_footnote(
      footnote = "The largest airports in each EU country (including the UK)",
      locations = cells_title(groups = c("title"))) %>% 
    opt_align_table_header(align = "left")

    ```
    

Value 4
:   
    ```{r, aviation-air-pollution-cost-aircraft}
    library(gt)
    library(magrittr)
    library(dplyr)
    library(tibble)
    tribble(
      ~ type, ~ distance, ~ emission_class, ~ aircraft_type, ~ eur_lto, ~ eur_pax_km, ~ eur_pax,
      "Short-haul", 500, "Low", "Bombardier CRJ900", 106, 0.29, 1.49,
      "Short-haul", 500, "High", "Embraer 170", 144, 0.32, 1.60,

      "Medium- haul", 1500, "Low", "Airbus 320", 174, 0.07, 1.17,
      "Medium- haul", 1500, "High", "Boeing 737", 195, 0.12, 1.66,
      "Medium- haul", 3000, "Low", "Airbus 320", 230, 0.05, 1.55,
      "Medium- haul", 3000, "High", "Boeing 737", 258, 0.07, 2.20,
      
      "Long-haul", 5000, "Low", "Airbus 340", 528, 0.03, 1.79,
      "Long-haul", 5000, "High", "Boeing 777", 876, 0.04, 2.02,
      "Long-haul", 15000, "Low", "Airbus 340", 748, 0.02, 2.53,
      "Long-haul", 15000, "High", "Boeing 777", 1.240, 0.02, 2.86
      ) %>% 
      dplyr::group_by(type) %>% 
      gt() %>%
      cols_label(
        type = "Type of flight",
        distance = "Distance [km]",
        emission_class = "Emission class",
        aircraft_type = "Example of aircraft type",
        eur_lto = "€ per LTO",
        eur_pax_km = "€-cent per pax km",
        eur_pax = "€ per pax"
    ) %>% 
    tab_header(title = md("Marginal air pollution costs of aviation for selected cases")) %>%
    tab_footnote(
      footnote = "For the following emissions: NH_3, NMVOC, SO2, NOx, PM2.5 and PM10",
      locations = cells_title(groups = c("title"))) %>% 
    tab_footnote(
      footnote = "For the cost factors for air pollution costs, the emissions
                  during the LTO cycle are mainly relevant, as the cruise
                  emissions lead to almost no damage costs.",
      locations = cells_title(groups = c("title"))) %>% 
    tab_footnote(
      footnote = "adjusted from € 2016 to € 2019 prices",
      locations = cells_column_labels(starts_with("eur"))) %>% 
    opt_align_table_header(align = "left")

    ```

Sources 1 to 4
:   “Handbook on the external costs of transport”, CE Delft,
    January 2019 (commissioned by European Union DG Move),
    tables 24, 49, 17 and 23
    https://www.cedelft.eu/en/publications/2311/handbook-on-the-external-costs-of-transport-version-2019
    <br>Also available from the Publications Office of the EU.
    
## Description

The CO2 price used to calculate the external costs of climate change
in **Value 1** is based on the cost avoidance approach.
Extract from the “Handbook on the external costs of transport”: 
*It determines external cost valuation factors (i.e. shadow prices) by
determining the cost to achieve a particular policy target
(e.g. EU CO2 reduction targets).
This is done by estimating an avoidance cost function, which provides a proxy
for the supply of environmental quality.
It determines how much it would cost to supply an additional level of
environmental quality (e.g. reduction of one additional tonne of CO2).
Based on this cost curve, the minimal cost required to meet the policy target
is estimated.
The assumption is that this policy target reflects collective preferences with
respect to the externality concerned and hence, that the minimum cost to reach
this target is a good proxy of the (collective) willingness-to-pay (WTP)
to avoid the damage caused by the externality.*

The calculations for **Value 2 and 3** have been made on the basis of
the cost factor of €100 per tonne of $CO_2$ equivalent, the central value
for short- and medium-run estimates given in value 1.
As was the case in the 2008 Handbook, it is further assumed that the aviation
emissions relevant for air quality are restricted to the emissions
in the landing and take-off (LTO) phases.
The total emissions have been cross-checked with the total emission database
from the European Monitoring and Evaluation Programme under the auspices of
the European Environmental Agency (EMEP/EEA).
The values are for passenger flights.


## Further reading

### EASA/EEA/EUROCONTROL

* European Environmental Report, 2019<br>
  https://www.eurocontrol.int/publication/european-aviation-environmental-report-2019 <br>
  The document provides an updated assessment of the environmental performance
  of the aviation sector published in the first report of 2016.
  The continued growth of the sector has produced economic benefits and
  connectivity within Europe and is stimulating investment in novel technology.
  This draws on a wider pool of expertise and innovative approaches from
  other sectors, thereby creating potential new opportunities to address
  the environmental impacts of aviation.
  However, it is recognised that the contribution of aviation activities to
  climate change, noise and air quality impacts is increasing, thereby
  affecting the health and quality of life of European citizens.
  
  Para. 5.3 touches on environmental charges levied by some airports.


* European Commission <br>
  https://ec.europa.eu/clima/policies/ets/auctioning_en <br>
  https://www.eex.com/en/
  
  Extracted from the EC website: “*Twenty-eight countries (25 EU Member States
  and 3 EEA/EFTA countries) auction their allowances on the common auction
  platform.
  The common auction platform is nominated for up to five years by a joint
  procurement between the Commission and the participating countries,
  in accordance to the rules laid down by the joint procurement agreement.
  Currently, the European Energy Exchange (EEX) in Leipzig is the common
  auction platform.
  Some countries participating in the EU ETS have opted out of the common
  auctioning platform: Germany has nominated EEX as its opt-out platform.
  Poland is making use of the common auction platform to auction its share of
  allowances until the appointment of its opt-out platform.
  ICE Futures Europe (ICE) in London acts as the United Kingdom's platform.
  The Withdrawal Agreement between the EU and the United Kingdom ensures that UK
  operators remain subject to compliance obligations for the years
  2019 and 2020.*”

* ECAC (2011)
  NOx emission classification scheme, ECAC recommendation ECAC27/4, 11 September 2011 <br> 
  https://www.ecac-ceac.org/documents/

* UK Department for Environment, Food and Rural Affairs (DEFRA)
  <br>Air quality damage cost update 2019, prepared by Ricardo Energy & Environment
  (Ricardo-AEA Ltd) https://uk-air.defra.gov.uk/assets/documents/reports/cat09/1902271109_Damage_cost_update_2018_FINAL_Issue_2_publication.pdf
  
  ```{r, pm-nox-damage-cost-estimate}
  library(tibble)
  library(gt)
  
  tribble(
    ~a , ~b , ~c , ~e ,
    "PM2.5", "194 (222)", "41 (46)", "560 (639)",
    "NOx", "12 (13)", "1.1 (1.2)", "45 (51)"
  ) %>% 
    gt() %>% 
    tab_header(
      title = "Revised sector PM and NOx damage cost estimates and sensitivity boundaries",
      subtitle = "(2017 prices, impacts discounted to 2017, average exchange rate 2017)"
    ) %>%
    tab_spanner(
      label = "damage cost sensitivity range",
      columns = vars(c, e)
    ) %>% 
    cols_label(
      a = "Emission emitted by aircraft",
      b = md("Central damage cost<br>£(€)/kg"),
      c = md("Low<br>£(€)/kg"),
      e = md("High<br>£(€)/kg")
    ) %>% 
    tab_footnote(
      footnote = "PM2.5 is the preferred metric for the change in PM emissions.",
      locations = cells_column_labels(columns = vars(a))
    ) %>% 
    tab_source_note(
      source_note = "Extract from tables 20-21 of the 'Air quality damage cost update 2019'"
    ) %>% 
    opt_align_table_header(align = "left")
  
  ```


  

## Related standard inputs

[Amount of emissions released by fuel burn] and [rate of fuel burn].


## Comments

With regard to NOx charges, a LTO NOx charge is currently applied at several
European airports.
The level of the charge per kg of NOx is set according to the cost of damage
caused by NOx to local air quality (LAQ), at or near airports.
The charge is levied in several countries, namely Sweden, the UK, Germany,
Demark and Switzerland.
Two examples are provided below.

### London Heathrow
At London Heathrow, a NOx emission charge is payable on each movement
(per LTO cycle) by a fixed-wing aircraft over 8 618 kg.
The charge per kg of NOx is calculated on the aircraft’s ascertained NOx
emission[^An aircraft’s ascertained NOx emission means the product of
the engine NOx emission as set out in the ICAO Emission Database and
based on the number of engines on the aircraft.].<br>
https://www.heathrow.com/content/dam/heathrow/web/common/documents/company/doing- business-with-heathrow/flights-condition-of-use/structure-of-charges- decision/Airport_Charges_Decision-5-August-2015.pdf

The Emissions Airport Charges for 2020 **per kg of NOx is £16.84**
(€20.5 at the 2019 exchange rate) “Decision – 2020 Airport Charges”<br> https://www.heathrow.com/content/dam/heathrow/web/common/documents/company/doing-business-with-heathrow/flights-condition-of-use/conditions-of-use-documents/Heathrow_Airport_Charges_and_Conditions_of_Use_2020.pdf


### Copenhagen Airports

Copenhagen Airports support, through financial incentives, the use of engine
types which emit the lowest emissions.
The model used for calculating emissions and charges is based on the models
currently found in Sweden and Switzerland.
The model works with absolute NOx emissions of a specific aircraft engine
during a standardised landing and take-off cycle (LTO).
NOx emissions per aircraft engine are based on the ICAO specifications and
guidelines.
The published emissions charge (November 2020) is **kr.16.50 per kg NOx**
(€2.2 at the 2019 exchange rate).<br>
https://www.cph.dk/en/cph-business/aviation/charges-and-slot/calculation-of-emission-charges

