# Fleet CNS capability {#sec-fleet-cns-capability}

## EUROCONTROL recommended values

The subsections below provide statistics on flights and aircraft with certain 
Communication, Navigation and Surveillance (CNS) equipment and capabilities.

```{r}
library(readxl)
library(dplyr)
library(gt)
```

### Navigation (NAV) flight capabilities

The values presented in the tables hereafter are weighted by flight and represent
the values for 2019.

```{r}
#| label: tbl-nav-flight
#| tbl-cap: NAV flight capabilities as a % of total flights

read_xlsx("data/cns_capability.xlsx",
          sheet = "nav_total") %>% 
  gt() %>% 
  cols_label(
    capability = "Capability",
    flights = "Number of flights",
    share_flights = "As a % of all flights"
  ) %>% 
  fmt_number(
    columns = flights,
    decimals = 0
  ) %>% 
  fmt_percent(
    columns = share_flights,
    decimals = 1
  ) %>% 
  cols_align(
    align = "right",
    columns = flights
  ) %>% 
  cols_align(
    align = "center",
    columns = share_flights
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(33),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "The number of flights is number of arrivals or departures, or arrivals and departures in line with the selection criteria in the CNS dashboard",
    locations = cells_column_labels(
      columns = flights
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-nav-flight_top10
#| tbl-cap: NAV flight capabilities as a % at the top 10 European Airports

read_xlsx("data/cns_capability.xlsx",
          sheet = "nav_top10") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    gbas = "A - GBAS landing system",
    lpv = "B - LPV (APV with SBAS)",
    rnav = "D - RNAV 1",
    rnp = "O - Basic RNP 1",
    rnp_apch = "S - RNP APCH"
  ) %>% 
  fmt_percent(
    columns = gbas:rnp_apch,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = gbas:rnp_apch
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    airport ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "The values for GBAS are corrected values. They exclude DHC8 equipment only compatible with a GBAS precursor system",
    locations = cells_column_labels(
      columns = gbas
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-nav-flight_month
#| tbl-cap: NAV aircraft capability over a one-month period, July 2019

read_xlsx("data/cns_capability.xlsx",
          sheet = "nav_month") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    gbas = "A - GBAS landing system",
    lpv = "B - LPV (APV with SBAS)",
    rnav = "D - RNAV 1",
    rnp = "O - Basic RNP 1",
    rnp_apch = "S - RNP APCH"
  ) %>% 
  fmt_percent(
    columns = gbas:rnp_apch,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = gbas:rnp_apch
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    airport ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-nav-flight_aircraft
#| tbl-cap: NAV flight capability by main aircraft model

read_xlsx("data/cns_capability.xlsx",
          sheet = "nav_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft",
    gbas = "A - GBAS landing system",
    lpv = "B - LPV (APV with SBAS)",
    rnav = "D - RNAV 1",
    rnp = "O - Basic RNP 1",
    rnp_apch = "S - RNP APCH"
  ) %>% 
  fmt_percent(
    columns = gbas:rnp_apch,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = gbas:rnp_apch
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    aircraft ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "The values for A319, A320, A330 and B737 aircraft (0.00%) are corrected values, as no such capability exist for these aircraft",
    locations = cells_column_labels(
      columns = lpv
    )
  ) %>% 
  tab_footnote(
    footnote = "This percentage represent the equipage rate of a GBAS precursor system called S-CATI",
    locations = cells_body(
      columns = gbas,
      rows = 5
    )
    ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

### Communication (COM) flight capabilities

The values presented in the tables hereafter are weighted by flight and represent
values for year 2019.

```{r}
#| label: tbl-com-flight
#| tbl-cap: COM flight capabilities as a % of total flights

read_xlsx("data/cns_capability.xlsx",
          sheet = "com_total") %>% 
  gt() %>% 
  cols_label(
    capability = "Capability",
    flights = "Number of flights",
    share_flights = "As a % of all flights"
  ) %>% 
  tab_row_group(
    label = "Voice",
    rows = 7:13
  ) %>% 
  tab_row_group(
    label = "Datalink",
    rows = 1:6
  ) %>% 
  tab_options(
    row_group.background.color = "#F5F7F9",
    row_group.font.weight = "bold"
  ) %>% 
  fmt_number(
    columns = flights,
    decimals = 0
  ) %>% 
  fmt_percent(
    columns = share_flights,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = share_flights
  ) %>% 
  cols_align(
    align = "right",
    columns = flights
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(30),
    capability ~ pct(40)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-com-flight_top10
#| tbl-cap: COM datalink flight capabilities as a % at the top 10 European airports

read_xlsx("data/cns_capability.xlsx",
          sheet = "com_top10") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    fis_acars = "E2 - D-FIS ACARS",
    pdc_acars = "E3 - PDC ACARS",
    cpdlc_atn_cdl = "J1 - CPDLC ATN VDL Mode 2",
    cpdlc_fans_vld = "J4 - CPDLC FANS 1/A VDL Mode 2",
    inmarsat = "J5 - CPDLC FANS 1/A SATCOM (INMARSAT)",
    iridium = "J7 - CPDLC FANS 1/A SATCOM (IRIDIUM)"
  ) %>% 
  fmt_percent(
    columns = fis_acars:iridium,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = fis_acars:iridium
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(8),
    airport ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-com-voice-top10
#| tbl-cap: COM voice flight capabilities as a % at the top 10 European airports

read_xlsx("data/cns_capability.xlsx",
          sheet = "com_voice_top10") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    hf_rtf = "H – HF RTF",
    inmarsat = "M1 – ATC RTF SATCOM (INMARSAT)",
    mtsat = "M2 – ATC RTF (MTSAT)",
    iridium = "M3 – ATC RTF (IRIDIUM)",
    uhf_rtf = "U – UHF RTF",
    vhf_rtf = "V – VHF RTF",
    vhf_spacing = "Y – VHF with 8.33 kHz channel spacing"
  ) %>% 
  fmt_percent(
    columns = hf_rtf:vhf_spacing,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = hf_rtf:vhf_spacing
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(9),
    airport ~ pct(37)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```
 
```{r}
#| label: tbl-com-aircraft
#| tbl-cap: COM datalink flight capabilities for main aircraft models

read_xlsx("data/cns_capability.xlsx",
          sheet = "com_datalink_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft",
    fis_acars = "E2 - D-FIS ACARS",
    pdc_acars = "E3 - PDC ACARS",
    cpdlc_atn_cdl = "J1 - CPDLC ATN VDL Mode 2",
    cpdlc_fans_vld = "J4 - CPDLC FANS 1/A VDL Mode 2",
    inmarsat = "J5 - CPDLC FANS 1/A SATCOM (INMARSAT)",
    iridium = "J7 - CPDLC FANS 1/A SATCOM (IRIDIUM)"
  ) %>% 
  fmt_percent(
    columns = fis_acars:iridium,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = fis_acars:iridium
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(14)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```
 
```{r}
#| label: tbl-com-voice-aircraft
#| tbl-cap: COM voice flight capabilities for main aircraft models

read_xlsx("data/cns_capability.xlsx",
          sheet = "com_voice_aircraft") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    hf_rtf = "H – HF RTF",
    inmarsat = "M1 – ATC RTF SATCOM (INMARSAT)",
    mtsat = "M2 – ATC RTF (MTSAT)",
    iridium = "M3 – ATC RTF (IRIDIUM)",
    uhf_rtf = "U – UHF RTF",
    vhf_rtf = "V – VHF RTF",
    vhf_spacing = "Y – VHF with 8.33 kHz channel spacing"
  ) %>% 
  fmt_percent(
    columns = hf_rtf:vhf_spacing,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = hf_rtf:vhf_spacing
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(9),
    airport ~ pct(37)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

### Surveillance (SUR) flight capabilities

The values presented in the tables hereafter are weighted by flight and represent
the values for 2022.

```{r}
#| label: tbl-sur-flight
#| tbl-cap: SUR flight capabilities as a % of total flights

read_xlsx("data/cns_capability.xlsx",
          sheet = "sur_total") %>% 
  gt() %>% 
  cols_label(
    capability = "Capability",
    flights = "Number of flights",
    share_flights = "As a % of all flights"
  ) %>% 
  tab_row_group(
    label = "ADS-B",
    rows = 10:15
  ) %>% 
  tab_row_group(
    label = "SSR Mode S",
    rows = 1:9
  ) %>% 
  tab_options(
    row_group.background.color = "#F5F7F9",
    row_group.font.weight = "bold"
  ) %>% 
  fmt_number(
    columns = flights,
    decimals = 0
  ) %>% 
  fmt_percent(
    columns = share_flights,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = share_flights
  ) %>% 
  cols_align(
    align = "right",
    columns = flights
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    capability ~ pct(70),
    flights ~ pct(15),
    share_flights ~ pct(15)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "Figures in this value are per flight from departure to destination, while values per airport are per flight but also include any flight departing from or arriving at that airport",
    locations = cells_column_labels(
      columns = flights
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-ssr-top10
#| tbl-cap: SSR/Mode S declared capabilities in flight plans at the top 10 European airports

read_xlsx("data/cns_capability.xlsx",
          sheet = "ssr_top10") %>% 
  gt() %>% 
  tab_spanner(
    "Capability",
    columns = L:A
  ) %>% 
  fmt_percent(
    columns = L:A,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = L:A
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    L:A ~ pct(9),
    Airport ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-adsb-top10
#| tbl-cap: ADS-B declared capabilities in flight plans at the top 10 European airports

read_xlsx("data/cns_capability.xlsx",
          sheet = "adsb_top10") %>% 
  gt() %>% 
  tab_spanner(
    "Capability",
    columns = B1:V2
  ) %>% 
  fmt_percent(
    columns = B1:V2,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = B1:V2
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    B1:V2 ~ pct(14),
    Airport ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-ssr-aircraft
#| tbl-cap: SSR/Mode S declared capabilities in flight plans for main aircraft models

read_xlsx("data/cns_capability.xlsx",
          sheet = "ssr_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft type"
  ) %>% 
  tab_spanner(
    "Capability",
    columns = L:A
  ) %>% 
  fmt_percent(
    columns = L:A,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = L:A
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    L:A ~ pct(9),
    aircraft ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

```{r}
#| label: tbl-adsb-aircraft
#| tbl-cap: ADS-B declared capabilities in flight plans for main aircraft models

read_xlsx("data/cns_capability.xlsx",
          sheet = "adsb_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft type"
  ) %>% 
  tab_spanner(
    "Capability",
    columns = B1:V2
  ) %>% 
  fmt_percent(
    columns = B1:V2,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = B1:V2
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    B1:V2 ~ pct(14),
    aircraft ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL Communication, Navigation & Surveillance (CNS) Dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)*")
  )
```

## Description

The CNS dashboard provides information for monitoring fleet capabilities and 
preparing performance-based navigation (PBN) deployment plans. It does so by 
analysing CNS and PBN information contained in ICAO flight plans, and generates
reports on aircraft or flight characteristics. Note that the capability of 
aircraft which do not submit flight plans is not covered in the above figures. 
The missing information to a large extent for general aviation (GA) flights.

The tool provides statistics on equipment and capability such as:

* Communication: FMC WPR ACARS, HF RTF; CPDLC FANS 1/A HFDL; etc.

* Navigation: RNAV 5, RNAV 1, RNP 1, RNP APCH (including LPV capability), GBAS, etc.

* Surveillance: ADS-B, ADS-C, Mode S transponder, etc.

Different periods of time, airports, airlines or aircraft types (depending on 
the user profile) can be analysed. 

To access the dashboard, you first need to register on the OneSky Online portal
using the link in the source above.

## Comments

The numbers of flights and aircraft provided by the CNS dashboard are derived 
from flight plans submitted to the EUROCONTROL Network Manager (NM). Consequently,
the statistics do not include the capability of aircraft flying in uncontrolled 
airspace or under VFR and thus do not submit flight plans to the NM.

On-board capability and equipment data made available via the CNS dashboard are
those declared in ICAO FPLs by operators. The information is therefore only as 
reliable as declared. For detailed analysis, additional local assessment is recommended. 

## Related inputs

@sec-number-of-ifr-flights [Number of IFR flights](#sec-number-of-ifr-flights)

@sec-cns-infrastructure [CNS infrastructure](#sec-cns-infrastructure)

@sec-pbn-and-precision-approach-procedures [PBN and precision approach procedure](#sec-pbn-and-precision-approach-procedures)

