# IFR average flight distance and flight duration {#sec-if-average-flight-distance-and-flight-duration}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

The table below presents an overview of the average flight distance in ECAC region,
as well as the average flight time in the same region for the period 2017-2021.

The data below was obtained by dividing the total distance actually flown and 
the total yearly IFR flight hours respectively by the yearly number of IFR
flights in the ECAC airspace.

```{r}
#| label: tbl-avg-flight-distance-duration
#| tbl-cap: Average IFR flight distance and duration in ECAC

read_xlsx("data/flight_distance_duration.xlsx") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    distancekm = "Distance in km",
    distancenm = "Distance in NM",
    time = "Flight time in min"
  ) %>% 
  fmt_number(
    columns = c(distancekm),
    use_seps = TRUE,
    decimals = 0
  )
```
*Source: EUROCONTROL, "Performance Review Report (PRR) 2021", 2022[@ectrlprr2022]; EUROCONTROL, "Performance Review Report (PRR) 2020", 2021[@ectrlprr2021]; STATFOR Interactive Dashboard[@ectrl:statfor:sid]*

:::{.callout-note}
The numbers for 2020 and 2021 are considerably lower 
due to the effect of pandemic.
:::

The values above are obtained by dividing the total distance actually flown and 
the total yearly IFR flight hours respectively by the yearly number of IFR flights
in ECAC airspace. The calculation stops at the borders of ECAC airspace.

With regard to flight distance, @fig-traffic-demand illustrates that nearly 90% of
IFR flight distances for flights arriving or departing within the Network Manager 
area (excl. overflights) are less than 1,000NM long.

![Distribution of traffic demand per distance band in 2022 within the EUROCONTROL Network Manager area](figures/traffic_demand_distribution.svg){#fig-traffic-demand}

## Related inputs

* @sec-ifr-flight-information-per-market-segment [IFR flight information per market segment](#sec-ifr-flight-information-per-market-segment)

* @sec-distance-flown-by-charging-zone [DIstance flown by charging zone](#sec-distance-flown-by-charging-zone)

* @sec-taxiing-times [Taxiing times](#sec-taxiing-times)

## References