# Load factor – passengers {#sec-load-factor-passengers}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

The load factor represents the percentage of seats filled by fare paying passengers
on a flight.

@tbl-load-pax-statfor shows the estimated passenger load factor according to EUROCONTROL
STATFOR values. These values were obtained by dividing the total number of 
passengers by the total number of available seats on the flights. This information 
is based on the data produced by Eurostat and covers the EU27+UK, and four EFTA
states.^[European Free Trade Association: Iceland, Liechtenstein, Norway and Switzerland]

```{r}
#| label: tbl-load-pax-statfor
#| tbl-cap: Passenger load factor – EUROCONTROL values

read_xlsx("data/load_factor_pax.xlsx",
          sheet = "statfor",
          range = "A1:B6") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    factor = "Load factor"
  ) %>% 
   cols_align(
    align = "center",
    columns = factor
  ) %>%
  cols_align(
    align = "left",
    columns = c(year)
  ) %>% 
  fmt_percent(
    columns = factor,
    decimals = 2
  ) %>% 
  tab_options(
    table.width = pct(70)
   ) %>% 
   cols_width(
    everything() ~ pct(50),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "year"
    )
  ) %>% 
   tab_source_note(
    source_note = md("*Source: [EUROCONTROL STATFOR Interactive Dashboard (PAX+)](https://www.eurocontrol.int/dashboard/statfor-interactive-dashboard)*")
  )
```

@tbl-load-pax-iata presents the evolution in the passenger load factor according to IATA data.
These values represent the ratio of revenue passenger km to available seat km.
The difference from values in Table 1 is the geographical coverage: IATA’s Europe
area is larger than EU Europe statistical area, it also covers countries such as 
Russia, Tajikistan, Turkmenistan and Uzbekistan. 

```{r}
#| label: tbl-load-pax-iata
#| tbl-cap: Passenger load factor – IATA values

read_xlsx("data/load_factor_pax.xlsx",
          sheet = "iata") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    factor = "Load factor"
  ) %>% 
   cols_align(
    align = "center",
    columns = factor
  ) %>%
  cols_align(
    align = "left",
    columns = c(year)
  ) %>% 
  fmt_percent(
    columns = factor,
    decimals = 2
  ) %>% 
  tab_options(
    table.width = pct(70)
   ) %>% 
   cols_width(
    everything() ~ pct(50),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "year"
    )
  ) %>% 
  tab_footnote(
    footnote = "The values used in this table come from IATA monthly Air Passenger Market Analysis reports for the month of December of each year. The data is extracted from the table “Air passenger market detail”, looking at Total market in Europe, column PLF(level) for calendar year.",
    locations = cells_column_labels(
      columns = factor
    )
  ) %>% 
   tab_source_note(
    source_note = md("*Source: IATA – Economics, Air Passenger Market Analysis for December 2017-2022. Available at [IATA - Economics](https://www.iata.org/en/publications/economics/?page=14&Search=Air%20Passenger%20Market%20Analysis%20-%20December%20&Year=2023&Year=2022&Year=2021&Year=2020&Ordering=DateDesc)*")
  )
```

A wide range of economic reports from IATA can be accessed on [IATA website](https://www.iata.org/en/publications/economics)
