---
editor: 
  markdown: 
    wrap: 72
---

# Passenger value of time {#sec-passenger-value-of-time}
```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

This input provides an estimation of the average value of passenger time
spent travelling, which might alternatively be spent on other activities
(e.g. working or leisure). It is essentially the **opportunity cost which corresponds to the monetary value associated with a passenger during a journey.** It shows how much a passenger would be willing to pay in order
to save time during a journey (e.g. by travelling on a quicker service
or using a faster transport mode), or how much 'compensation' they would
accept, directly or indirectly, for lost time lost.

In tables hereafter is presented a collection of the different values that can be used
for this purpose. It is to be noted that, in this section, the value of
time is not cited as a function of delay duration. This is an important
consideration when using the value. In this case, the longer the delay
duration, the higher the value.

```{r}
#| label: tbl-pas-time-value-eu
#| tbl-cap: Estimated value of travel time - EU25
read_xlsx("data/passenger_time_value.xlsx",
          sheet = "eu25") %>% 
  gt() %>% 
  cols_label(
    purpose = "Travel purpose",
    value = "Time value (€/hour)"
  ) %>% 
  fmt_currency(
   columns = c(value),
   currency = "EUR",
   decimals = 2
  ) %>% 
  tab_footnote(
  footnote = "Please note that the values above are adjusted to 2022 prices from 2002 based on inflation",
locations = cells_column_labels(
  columns = value
)
) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [European Commission,  “HEATCO, Developing Harmonised European Approaches for Transport Costing and Project Assessment – Deliverable 5 Proposal for Harmonised Guidelines”, 2006](https://trimis.ec.europa.eu/sites/default/files/project/documents/20130122_113653_88902_HEATCO_D5_summary.pdf)*")
  )
```

```{r}
#| label: tbl-pas-time-value-fr
#| tbl-cap: Estimated value of travel time - France
read_xlsx("data/passenger_time_value.xlsx",
          sheet = "france") %>% 
  gt() %>% 
  cols_label(
    purpose = "Travel purpose",
    value = "Time value (€/hour)"
  ) %>% 
  fmt_currency(
   columns = c(value),
   currency = "EUR",
   decimals = 2
  ) %>% 
  
tab_footnote(
  footnote = "Please note that the values above are adjusted to 2022 prices from 2015 based on inflation",
locations = cells_column_labels(
  columns = value
)
) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [French Ministry of Ecological Transition, Gouvernement Français, “Recommended values for calculating average long-distance travel”, 2019](https://www.ecologie.gouv.fr/sites/default/files/V.3.pdf)*")
  )
  
```

```{r}
#| label: tbl-pas-time-value-uk
#| tbl-cap: Estimated value of travel time - United Kingdom
read_xlsx("data/passenger_time_value.xlsx",
          sheet = "uk") %>% 
  gt() %>% 
  cols_label(
    purpose = "Travel purpose",
    value = "Time value (€/hour)"
  ) %>% 
  fmt_currency(
   columns = c(value),
   currency = "EUR",
   decimals = 2
  ) %>% 
  
tab_footnote(
  footnote = "Please note that the values above are adjusted to 2022 prices from 2014 based on inflation and using the exchange rate shown in general parameters",
locations = cells_column_labels(
  columns = value
)
) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [Airports Commission, “Values of time used to estimate passenger delay costs in the UK airport system”, 2014](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/372606/AC08a_tagged.pdf)*")
  )
  
```

## When to use the input?

This input is expected to become useful in any study that looks at the 
opportunity cost of the use of air transport, delays, cancellations, etc. 
It provides a perspective on the impact that a change in the air transport
can have on the passenger.

## Comments

When looking into the values in the table, a few points regarding the sources 
should be taken into consideration:

* The source used for the UK values sets out a methodology for analysis which 
has been undertaken to estimate benefits from reduced delay time to airlines 
and passengers from changes in aviation capacity constraints in the UK for 11 
airports. 

* The values for France rely on a working paper on recommended values for
calculating the components of a socio-economic net present value, which include 
travel time. The assessment therefore covers social, environmental and economic 
effects.

* Regarding the numbers for EU25, they remain a reference if a European value is
sought. The objective of the study from which they were derived is to propose 
harmonised guidelines for project assessment for transnational projects in
Europe. It provides monetary estimates for the values of time saved for 
employer’s business and for passenger non-work trips (e.g. commuting, shopping
and leisure).

## Other values to consider

Some additional values are available to EUROCONTROL, which, although not 
constituting the perceived key inputs, may be useful for specific purposes of 
the user of these inputs. These values are presented below.

***Value of time of a business aviation passenger***

**€153 per hour.** This value was provided to EUROCONTROL by airline members of 
the SESAR CBA team in 2012 and is, therefore, adjusted from 2011 prices. 

***Value of passenger time in the US for high-speed rail passengers***

Based on US DOT guidance on passenger value of time for air and high-speed rail 
travel by purpose of trip. The numbers are adjusted to inflation from 2015 prices
and using the exchange rate presented in @tbl-eurostat-exchange-rates.

## Further reading

Below are listed some additional sources recommended for consultation:

* International Transport Forum (ITF), “What is the Value of Saving Travel Time?” Feb. 2019[@itf:2019]

* Economic Development Research Group Inc., “Passenger Value of Time, Benefit-Cost Analysis and Airport Capital Investment Decisions,” 2015[@passtimevalue2015]

* University of Leeds, “Values of travel time in Europe: Review and meta-analysis,” 2016[@leedsuniversity2016]

* University of Leeds, “European Wide-Meta Analysis of Values of Travel Time,” May 2012[@leedsuniversity2012]

## References