# PBN and precision approach procedures {#sec-pbn-and-precision-approach-procedures}

```{r}
library(readxl)
library(dplyr)
library(gt)
```


## EUROCONTROL recommended values

The values hereafter present the proportion and list of airports and runway ends 
in ECAC with published Performance-Based Navigation (PBN) instrument approach 
procedures.[@pbn:tool]

```{r}
#| label: tbl-pbn-deployment
#| tbl-cap: PBN approach deployment status - September 2020

read_xlsx("data/pbn_procedures.xlsx",
          sheet = "pbn_deployment") %>% 
  gt() %>% 
  cols_label(
    approach = "Approach type",
    runway_nb = "Runway ends covered (Nb)",
    runway_share = "Runway ends covered (%)",
    apt_nb = "Airports covered (Nb)",
    apt_share = "Airports covered (%)"
  ) %>% 
  fmt_percent(
    columns = c(runway_share, apt_share),
    decimals = 1
  ) %>% 
  fmt_number(
    columns = c(runway_nb, apt_nb),
    decimals = 0
  ) %>% 
  cols_align(
    align = "center",
    columns = c(runway_share, apt_share)
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    runway_nb:apt_share ~ pct(15),
    approach ~ pct(40)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
tab_source_note(
  source_note = md("*Source: [EUROCONTROL PBN Approach Map Tool](https://www.eurocontrol.int/platform/performance-based-navigation-map-tool)*")
)
```

```{r}
#| label: tbl-ils-deployment
#| tbl-cap: ILS Cat I, Cat II/III deployment status - September 2020

read_xlsx("data/pbn_procedures.xlsx",
          sheet = "ils_deployment") %>% 
  gt() %>% 
  cols_label(
    approach = "Approach type",
    runway_nb = "Runway ends covered (Nb)",
    runway_share = "Runway ends covered (%)",
    apt_nb = "Airports covered (Nb)",
    apt_share = "Airports covered (%)"
  ) %>% 
  fmt_percent(
    columns = c(runway_share, apt_share),
    decimals = 1
  ) %>% 
  fmt_number(
    columns = c(runway_nb, apt_nb),
    decimals = 0
  ) %>% 
  cols_align(
    align = "center",
    columns = c(runway_share, apt_share)
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    runway_nb:apt_share ~ pct(15),
    approach ~ pct(40)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
tab_source_note(
  source_note = md("*Source: [EUROCONTROL PBN Approach Map Tool](https://www.eurocontrol.int/platform/performance-based-navigation-map-tool)*")
)
```

## Description

The EUROCONTROL PBN Approach Map tool illustrates the deployment of PBN 
instrument approach procedures against objectives set in ICAO Assembly Resolution 
37-11 and the European Regulation on PBN 
(in particular Commission Implementing Regulation (EU) 2018/1048 of 18 July 2018). 

PBN approaches include instrument approach procedures compliant with the following
navigation specifications of the PBN Manual (ICAO Doc 9613):

* RNP APCH

* RNP AR APCH

The PBN Approach Map tool provides a list of current and planned airports and 
runway ends covered by each type of approach. The tool gives information about:

* deployment progress since 2012, on the basis actual publications

* future deployment trends based on publication plans communicated to EUROCONTROL and ICAO

* the availability of PBN approaches with vertical guidance (APV) on all runway 
ends or on runway ends without precision landing (e.g. ILS, MLS or GBAS)

* the deployment status for ECAC, individual countries, PCP airports and individual airports

In September 2020 (on the basis of the AIRAC cycle), 1,576 runway ends were 
equipped with instrument approach procedures and 763 airports had instrument 
approach procedures.

## Comments

The PBN Approach Map tool is updated in accordance with publications with every 
AIRAC cycle. It therefore provides up-to-date information on the current deployment status.

Information about publication plans captured in the map is collected from 
individual countries and coordinated with the ICAO EUR/NAT regional office and
other implementation-funded programmes. If and when these publication plans 
materialise depends on a number of factors including:

* difficulties collecting obstacle data for procedure design

* unforeseen problems in the procedure design phase

* delays in approval for publication by the supervisory authority

Implementation plans should consequently not be considered to be a State’s commitment.

## References


