# Amount of emissions released by fuel burn {#sec-amount-of-emissions-released-by-fuel-burn}


```{r}
library(readxl)
library(dplyr)
library(gt)
library(tidyverse)
```

## EUROCONTROL recommended values 

This input represents the amount of emissions produced by combustion of aviation
fuel (i.e. Kerosene), focusing on the main types of pollutants.

```{r}
#| label: tbl-emissions-per-kg-fuel
#| tbl-cap: Estimated amount of emissions released per kg of fuel burnt

tribble(
  ~type, ~emissions,
  "CO2"        , "3.15 kg",
  "H2O"        , "1.237 kg",
  "SOx"        , "0.00084 kg",
  ""           , "0.000114 kg",
  "NOx"        , "0.0148 kg",
  "HC"         , "0.00032 kg",
  "CO"         , "0.00325 kg",
  "PM volatile", "0.000092 kg"
) %>% 
  gt() %>% 
  tab_header(title = "Emissions per kg of fuel burnt",
             subtitle = "fuel = Kerosene") %>%
  cols_align(
    align = "right",
    columns = "emissions"
  ) %>% 
  cols_label(
    type = md("**Emission type**"),
    emissions = md("**Emission**")
  ) %>%
  # tab_style(
  #   style = list(
  #     cell_text(weight = "bold")
  #   ),
  #   locations = cells_body(
  #     columns = "type"
  #   )
  # ) %>% 
  # tab_spanner(
  #   "Emissions per kg of fuel burnt",
  #   1:2
  # ) %>% 
  tab_footnote(
    footnote = "Source 1",
    locations = cells_body(columns = emissions, rows = 1:2),
    placement = "right"
  ) %>% 
  tab_footnote(
    footnote = "assuming 440 ppm, Source 2",
    locations = cells_body(columns = emissions, rows = 3),
      placement = "right"
  ) %>% 
  tab_footnote(
    footnote = "assuming 600 ppm, Source 2",
    locations = cells_body(columns = emissions, rows = 4),
      placement = "right"
  ) %>% 
  tab_footnote(
    footnote = "Source 3",
    locations = cells_body(columns = emissions, rows = 5:8),
    placement = "right"
  ) %>% 
  tab_options(
    table.width = pct(70),
    footnotes.marks = c(letters[1:10])
   )  %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(50),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source 1: [EUROCONTROL (2018) European Aviation Fuel Burn and Emissions Inventory System for the European Environment Agency](https://www.eurocontrol.int/publication/european-aviation-fuel-burn-and-emissions-inventory-system-feis-european-environment)*")
  ) %>% 
tab_source_note(
    source_note = md("*Source 2: [MIT (2010) Laboratory for aviation and the environment – Guidance on the use of AEDT for gridded aircraft emissions in atmospheric models](https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.719.2090&rep=rep1&type=pdf)*")
  ) %>% 
  tab_source_note(
    source_note = md("*Source 3: [EUROCONTROL, EASA, EU (2022) European Aviation Environmental Report, EU27+FTA aviation emissions for year 2021](https://www.eurocontrol.int/publication/european-aviation-environmental-report-2022)*")
  )
```

The Committee on Aviation Environmental Protection (CAEP), a technical committee
of the ICAO Council, recommends the use a conversion factor of **3.16 g of CO~2~per gram of Jet A**. 
The 3.16 value can be found in ICAO Doc 9889, 1st edition, 2011, and other documents.

However, in Europe, as early as 2009, Commission Decision 2009/339/EC indicated 
an **emission factor of 3.15 for the mass conversion from Jet A to CO~2~** for 
the period after January 2021.

In view of the above, emission factor 3.15 should continue to be used in SESAR 
activities, for the sake of internal consistency within the programme, unless 
the EU ETS decides to move to 3.16. Emission factor 3.16 should be used when the
evaluation concerns comparisons with studies carried out within the ICAO framework
or using the factor recommended by ICAO, in order to ensure external consistency.^[Digital European Sky guidance and expert opinion]

## Other possible sources

* *European Aviation Environmental Report Series* [@eaer2022]
  
  The 2022 edition shows the aviation sector has taken steps to address 
  environmental challenges but also that more decisive actions are needed. 
  In particular, the latest trends in aircraft engine emissions are shown in 
  section 3.2. of the document.

* *European Environment Agency (2023), EMEP^[EMEP is a cooperative programme for monitoring and evaluation of the long-range transmission of air pollutants in Europe]/EEA air pollutant emission inventory guidebook 2023* [@eea:2019]

  Provides values for emission factors and fuel consumption in different phases 
  of flight – including taxiing – for different aircraft types, using three 
  different levels of accuracy and complexity, in section 1.A.3.a of the document.

* *ICAO Aircraft Engine Emissions Databank*

  Available on the [EASA website](https://www.easa.europa.eu/en/domains/environment/icao-aircraft-engine-emissions-databank), 
  this source contains information on exhaust emissions of production aircraft engines,
  covering engine types whose emissions are regulated, namely turbojet and turbofan 
  engines with a static thrust greater than 26.7 kilonewtons.

* *Swiss Federal Office of Civil Aviation, Aircraft Engine Emissions* [@foca:aee]
   
   Presents a measurement and calculation methodology for aircraft piston engine
   emissions in order to improve aviation emission inventories, as developed by FOCA.

* *Swedish Defence Research Agency* [@envimpact]

  Holds a database of emission indices of NOx, HC, and CO, with corresponding 
  fuel flows for turboprop engines.

## When to use the input? 

This input is recommended for a wide use in assessments that focus on the 
assessment of environmental impact from the burning of fuel at any stage of the flight.

## Related inputs

@sec-rate-of-fuel-burn [Rate of fuel burn](#sec-rate-of-fuel-burn)

@sec-cost-of-emissions [Cost of emissions](#sec-cost-of-emissions)

@sec-ifr-flight-information-per-market-segment [IFR flight information per market segment](#sec-ifr-flight-information-per-market-segment)

## References
