# ANSP employment costs {#sec-ansp-employment-cost}

```{r}
library(readxl)
library(dplyr)
library(gt)
library(here)
```


## EUROCONTROL recommended values

This section considers ANSPs’ average annual employment costs for one Full Time 
Equivalent (FTE) in euros by category of staff. @tbl-ansp-employment-cost-general 
shows the recommended values.

```{r}
#| label: tbl-ansp-employment-cost-general
#| tbl-cap: Average yearly ANSP employment costs (2021 prices)

read_xlsx(here("data", "ansp_employment.xlsx"),
          sheet = "general") %>%
  gt()%>%
  cols_label(
    funct = "Staff function",
    ectrl_area = "EUROCONTROL area",
    ses_area = "SES area"
  )%>%
  fmt_currency(
    currency = "EUR",
    columns = ectrl_area:ses_area,
    decimals = 0
  )%>%
   tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = funct:ses_area,
      rows = 3
    )
   )%>%
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    funct ~ px(50),
    ectrl_area ~ px(25),
    ses_area ~ px(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source = md("*Source: [EUROCONTROL PRU, ‘ATM Cost-Effectiveness (ACE) Benchmarking Report (2023 Edition)’](https://ansperformance.eu/publications/prc/ace/)*")
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(33),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

One FTE is the equivalent of a single person carrying out a particular job or 
activity working on a full-time basis for a year. A part-time employee working
half-time would be counted as 0.5 FTEs. A full-time Air Traffic Control Officer
(ATCO) working two-thirds of their time on duty in ops and one third of their 
time on teaching at a training academy would be counted as a 0.67 FTE ATCO in 
ops and a 0.33 FTE Support Staff.

Employment costs comprise gross wages and salaries, payment for overtime, employer
contributions to social security schemes and taxes, pension contributions and 
other benefits. For a study on employment costs, the categories of staff working 
in an ANSP have been divided into two:

* ATCOs in ops: ATCOs participating in an activity which is either directly 
  related to the control of traffic or where there is a necessary requirement 
  for ATCOs to be able to control traffic
  
* Support staff or non-ATCO in ops: this category includes all other staff. It
  includes ATCOs on other duties (participating in an activity outside ops, such 
  as special projects, teaching at a training academy, providing instruction in 
  a simulator, working in a full-time management position, etc.), trainees, ATC
  assistants, technical and operational support staff, administration staff, and others
  
@tbl-ansp-empl-cost gives an overview of individual ANSPs and average European 
system FTE costs for the two categories. The values were calculated based on
information extracted from EUROCONTROL ACE Report.


```{r}
#| label: tbl-ansp-empl-cost
#| tbl-cap: Individual ANSP and average European system FTE costs in thousand euros (2021 prices)

read_xlsx(here::here("data", "ansp_employment.xlsx"),
          sheet = "all") %>%
  gt()%>%
  cols_label(
    ansp = "ANSP",
    atco_ops = "ATCO in ops",
    support = "Support staff",
    all = "All staff"
  )%>%
  fmt_currency(
    currency = "EUR",
    columns = atco_ops:all,
    decimals = 0
  )%>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = ansp:all,
      rows = 39
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    ansp ~ pct(40),
    atco_ops:all ~ pct(20)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source = md("*Source: [EUROCONTROL PRU, ‘ATM Cost-Effectiveness (ACE) Benchmarking Report (2023 Edition)’](https://ansperformance.eu/publications/prc/ace/)*")
  )
```

The values in @tbl-ansp-empl-cost were calculated using values provided in 
Annex 5, Tables 0.3 and 0.5 of the source document. The employment costs refer
to gate-to-gate cost (i.e. en-route and terminal costs) and are expressed in 2021 prices.

## Related inputs

@sec-atm-cost-effectiveness-indicators [ATM cost-effectiveness indicators](#sec-atm-cost-effectiveness-indicators)
