# ATM operational units {#sec-atm-operational-units}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

This value represents the number of ATC units (air traffic centres) providing ATC
services across Europe for the purpose of preventing collisions between aircraft,
and on the manoeuvring area between aircraft and obstructions, and expediting and 
maintaining an orderly flow of air traffic.

```{r}
#| label: tbl-atm-op-units
#| tbl-cap: Number of ATM operational units in Europe

read_xlsx(here::here("data", "atm_operational_units.xlsx"))%>% 
  gt() %>% 
  cols_label(
    year = "Year",
    ansp = "ANSP",
    acc = "ACC",
    app = "APP",
    twr = "TWR",
    afis = "AFIS units",
    atc = "ATC sectors"
  ) %>% 
  cols_align(
    align = "left",
    columns = c(year)
  ) %>% 
  cols_align(
    align = "right",
    columns = c(ansp:atc)
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "year"
    )
  ) %>% 
  tab_footnote(
    footnote = "ANSP - Air Navigation Service Provider",
    locations = cells_column_labels(
      column = ansp
    )
  ) %>% 
  tab_footnote(
    footnote = "ACC - Area Control Centre",
    locations = cells_column_labels(
      column = acc
    )
  ) %>% 
  tab_footnote(
    footnote = "APP - Approach Units",
    locations = cells_column_labels(
      column = app
    )
  ) %>% 
  tab_footnote(
    footnote = "TWR - Tower",
    locations = cells_column_labels(
      column = twr
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(14)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "AFIS - Airport/Aerodrome Flight Information Service",
    locations = cells_column_labels(
      column = afis
    )
  ) %>% 
  tab_footnote(
    footnote = "ATC - Air Traffic Control",
    locations = cells_column_labels(
      column = atc
    )
  ) %>% 
  tab_footnote(
    footnote = "Sakaeronavigatsia was added to the analysis in 2015",
    locations = cells_body(
      column = year,
      row = 5
    )
  ) %>% 
  tab_footnote(
    footnote = "In 2021, UkSATSE was not included in the analysis of operational units, while BHANSA was added to the list, impacting the numbers presented in the table",
    locations = cells_body(
      column = year,
      row = 11
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL ATM Cost-Effectiveness (ACE) Benchmarking Reports](https://www.eurocontrol.int/ACE/ACE-Home.html)*")) 
```

## Comment

Please note that the analysis presented in ACE Benchmarking reports, on which this
input was based, excludes elements related to services provided to military
operational air traffic (OAT), oceanic ANS, and landside airport management 
operations. It presents a review and comparison of ATM cost effectiveness for
the 38 ANSPs in Europe, which provide coverage for EUROCONTROL’s 41 Member States. 

## When to use this input?

The values presented above can be used in analyses where the level of granularity
of information per type of provider or per provider is important. This is typically
interesting for studies looking into costs or benefits per ANSP/ACC/Tower/other.

## Related inputs

@sec-atm-cost-effectiveness-indicators [ATM cost-effectiveness indicators](#sec-atm-cost-effectiveness-indicators)

