# CNS infrastructure {#sec-cns-infrastructure}

```{r}
library(readxl)
library(dplyr)
library(gt)
```


## EUROCONTROL recommended values

The values below show the number of civil-owned systems installed which are 
devoted to carrying out navigation and surveillance functions in ECAC.

### Navigation aids

@tbl-cns-ecac1 and @tbl-cns-ecac2 shows the number of Navigation aids in ECAC at 
the end of 2022. Please note there might be some marginal differences with the
latest available picture in the [EUROCONTROL Ground-based Navigation Infrastructure Map Tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)
(GNI tool) depending on the date of extraction.

```{r}
#| label: tbl-cns-ecac1
#| tbl-cap: Navigation aids in the ECAC Member States - part 1

read_xlsx(here::here("data", "cns_infrastructure.xlsx"),
          sheet = "navaids1") %>% 
  gt() %>% 
  cols_align(
    align = "right",
    columns = everything()
  ) %>% 
  cols_align(
    align = "left",
    columns = "State"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "State"
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(10)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 45
    )
  ) %>% 
  tab_footnote(
    footnote = "Considered in Serbia as service is provided by SMATSA",
    locations = cells_body(
      columns = State,
      rows = 28
    )
  ) %>% 
   tab_footnote(
    footnote = "Including Azores and Madeira",
    locations = cells_body(
      columns = State,
      rows = 33
    )
  ) %>% 
  tab_footnote(
    footnote = "Considering jointly the State of Serbia and the State of Montenegro",
    locations = cells_body(
      columns = State,
      rows = 36
    )
  ) %>% 
   tab_footnote(
    footnote = "Including Canary Islands; Ceuta and Melilla and Spain Continental",
    locations = cells_body(
      columns = State,
      rows = 39
    )
  ) %>% 
  tab_source_note (
    source = md("*[EUROCONTROL Ground-based Navigation Infrastructure Map Tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)*")
  )
```

```{r}
#| label: tbl-cns-ecac2
#| tbl-cap: Navigation aids in the ECAC Member States - part 2

read_xlsx(here::here("data", "cns_infrastructure.xlsx"),
          sheet = "navaids2") %>% 
  gt() %>% 
  cols_align(
    align = "right",
    columns = everything()
  ) %>% 
  cols_align(
    align = "left",
    columns = "State"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "State"
    )
  ) %>% 
  tab_spanner(
    "ILS",
    columns = 2:4
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(20)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 45
    )
  ) %>% 
  tab_footnote(
    footnote = "Considered in Serbia as service is provided by SMATSA",
    locations = cells_body(
      columns = State,
      rows = 28
    )
  ) %>% 
   tab_footnote(
    footnote = "Including Azores and Madeira",
    locations = cells_body(
      columns = State,
      rows = 33
    )
  ) %>% 
  tab_footnote(
    footnote = "Considering jointly the State of Serbia and the State of Montenegro",
    locations = cells_body(
      columns = State,
      rows = 36
    )
  ) %>% 
  tab_footnote(
    footnote = "Including Canary Islands; Ceuta and Melilla and Spain Continental",
    locations = cells_body(
      columns = State,
      rows = 39
    )
  ) %>% 
  tab_source_note (
    source = md("*[EUROCONTROL Ground-based Navigation Infrastructure Map Tool](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)*")
  )
```

The GNI Tool illustrates the current status, and the evolution plans for the
Ground-based Navigation Infrastructure in ECAC. The Navaids database is updated 
based on the information published in the [European AIS Database (EAD)](https://www.eurocontrol.int/service/european-ais-database) and the 
planning provided by Air Navigation Service Providers. Please note the evolution
plans shall not be considered as a commitment by a State or ANSP.

### Surveillance units

@tbl-cns-ecac-surv shows the number of Surveillance units in ECAC at the end of 2022.

```{r}
#| label: tbl-cns-ecac-surv
#| tbl-cap: Surveillance units in EUROCONTROL Member States in 2022

read_xlsx(here::here("data", "cns_infrastructure.xlsx"),
          sheet = "surveillance") %>% 
  gt() %>% 
  cols_align(
    align = "right",
    columns = everything()
  ) %>% 
  cols_align(
    align = "left",
    columns = "State"
  ) %>% 
    tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "State"
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    State ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "The WAM/ADS-B column lists the number of sensors. The configurations and system boundaries for several WAM/ADS-B implementations are complex and site specific. It is therefore not possible to consistently identify the corresponding number of systems",
    locations = cells_column_labels(
      column = 5
    )
    )%>% 
  tab_source_note (
    source = md("*Source: EUROCONTROL Surveillance Unit Surveillance Database; LSSIP+ Cycle 2022 – SUR Questionnaire*")
  )
```

The Mode-S, Primary Surveillance Radar (PSR) and Secondary Surveillance Radar 
(SSR) (i.e. Mode A/C) numbers are extracted from the surveillance database of 
the EUROCONTROL CNS unit.

According to the most recent figures, in Europe there are 526 Mode-S radars, 142
PSRs and 139 SSRs, either combined or standalone. As the allocation and 
implementation of Mode-S interrogator codes (ICs) require a coordinated approach,
every installation of a Mode-S radar is officially registered. The numbers of
PSRs and SSRs reported above are not necessarily accurate, as they are based on 
voluntary reports by the Member States on updates and changes to their surveillance 
infrastructure. Work on the collection of multilateration (MLAT)/ADS-B stations 
is still in progress.

The wide-area multilateration (WAM)/ADS-B (wide-area multilateration/automatic
dependent surveillance-broadcast) and ADS-B data originate from the database which
is maintained by the EUROCONTROL Surveillance and Code Coordination Unit and is
based on inputs from stakeholders. The ADS-B and WAM Section coordinates the 
deployment of initial ADS-B applications and WAM in Europe.

The WAM/ADS-B sensor count only includes sensors mainly used for surveillance of
airborne aircraft (e.g. in TMAs or en route). It does not include sensors mainly 
used for airport surface surveillance (e.g. airport MLAT used for A-SMGCS).

## Related inputs

@sec-fleet-cns-capability [Fleet CNS capability](#sec-fleet-cns-capability)

@sec-pbn-and-precision-approach-procedures [PBN and precision approach procedures](#sec-pbn-and-precision-approach-procedures)
