---
editor_options: 
  markdown: 
    wrap: 72
---

# General parameters {#sec-conversions .unnumbered}

The sections hereafter present some key values used in the Standard Inputs that 
can also serve as a reference for other uses.

## Inflation {#inflation .unnumbered}

The inflation levels have been calculated based on the Eurostat European Union 
region Harmonised Index of Consumer Prices (HICP)[@eurostat:HICP], as the yearly 
difference in the HICP. The HICP data was extracted from [Eurostat](http://ec.europa.eu/eurostat)
HICP - annual data (average index and rate of change) database
(prc_hicp_aind).The annual change in the index is shown in @tbl-inflation-table.

```{r}
#| label: tbl-inflation-table
#| tbl-cap: Annual average inflation values

library(readxl)
library(dplyr, quietly = TRUE)
library(gt)
library(here)

# read from CSV (exported directly from Eurostat, see R/export_inflation.R)
readxl::read_xlsx(here("data","values_2022.xlsx"),
                  sheet = "basic assumptions",
                  range = "C6:E34" 
                  ) %>%
  filter(2000 <= Year, Year <= 2023) %>%
  arrange(desc(Year)) %>% 
  # format the table
  gt() %>%
  cols_label(
    Year  = "Year",
    Index = "Index",
    Rate  = "Rate of change"
  ) %>%
  fmt_percent(
    columns = c(Rate),
    decimals = 1
  ) %>%
  fmt_number(
    columns = c(Index),
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = c(Year, Index, Rate)
  ) %>%
  tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
     Year:Rate ~ pct(30)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = Year
    )
  )
```

## Exchange Rate Conversion {#exchange-rates .unnumbered}

The currency exchange rates provided in @tbl-eurostat-exchange-rates are based 
on the European Central Bank (ECB) rates that are published daily. [@ECB:exchange_rates]. 
The rates correspond to the average rate for the year 2022 and are used across 
this document for the purpose of currency conversion.

```{r }
#| label: tbl-eurostat-exchange-rates
#| tbl-cap: Average euro foreign exchange rate in 2022
library(readr)
library(dplyr)
library(gt)

read_csv(here("data", "exchange_rate.csv"), show_col_types = FALSE) %>%
  gt() %>%
  cols_label(
    currency  = "Currency",
    to_eur    = "EUR/currency",
    from_eur  = "Currency/EUR"
  ) %>%
  cols_align(
    align = "center",
    columns = c(to_eur, from_eur)
  ) %>% 
  tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
     currency ~ pct(40),
     to_eur ~ pct(30),
     from_eur ~ pct(30)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )
```

## Cost of fuel {.unnumbered}

The cost of fuel used in this document is based on the 2022 average jet fuel
price provided by IATA,[@iata:JetFuelPrice], unless otherwise specified.
All conversions are done using the values specified on this page.

```{r}
#| label: tbl-cost-of-fuel
#| tbl-cap: Average jet fuel price in 2022
library(dplyr)
library(gt)

tribble(
  ~currency, ~price_barrel, ~price_gallon, ~price_kg,
  "USD", 136, 3.2, 1.1,
  "EUR", 129, 3.1, 1.0,
) %>%
  gt() %>%
  cols_label(
    currency = "Currency",
    price_barrel = "Price per barrel",
    price_gallon  = "Price per gallon",
    price_kg = "Price per kg"
  ) %>%
  cols_align(
    align = "center",
    columns = c(price_barrel, price_gallon, price_kg)
  ) %>% 
  tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
     currency ~ pct(40),
     price_barrel ~ pct(20),
     price_gallon ~ pct(20),
     price_kg ~ pct(20)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )
```

## Conversion values {.unnumbered}

```{r}
#| label: tbl-conversion-values
#| tbl-cap: Unit conversion values
tribble(
  ~from, ~to,
  "1 nautical mile (NM)", "1.852 km",
  "1 kilometre (km)"    , "0.53996 NM",
  "1 tonne (metric = 1 000 kg) of jet fuel", "325.33 US gallons",
  ""                                       , "1 235 litres",
  ""                                       , "7.8 barrels",
  "1 barrel (bbl) of jet fuel", "42 US gallons",
  ""                          , "158.99 litres",
  ""                          , "0.1291 tonne = 129.10 kg",
  "1 US gallon of jet fuel (US gal)", "3.7854 litres",
  ""                                , "3.073 kg",
  ""                                , "6.7764 lb",
  "Density of kerosene", "0.812 kg/litre",
  "1 litre of fuel (l)", "0.26417 US gallons",
  "1 kilogramme of fuel (kg)", "2.2046 lb",
  "1 pound of fuel (lb)", "0.45359 kg",
  "1 US ton", "1.01605 tonne"
) %>% 
  gt() %>%
  cols_label(
    from = "From",
    to = "To",
    ) %>% 
  tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
     from ~ pct(50),
     to ~ pct(50)
  ) %>%
  tab_options(
    column_labels.font.weight = "bold"
  )
```

## References
