# Cost of diversion {#sec-cost-of-diversions}

## EUROCONTROL recommended values

```{r}
library(readxl)
library(dplyr)
library(gt)
```

@tbl-diversion-comm presents the estimated **average cost of the diversion of a flight to an airport other than the one initially planned.** 
These values have been estimated through an extensive consultation process with the Airline members.

```{r}
#| label: tbl-diversion-comm
#| tbl-cap: Estimated average cost of a diversion for scheduled aviation

read_xlsx(here::here("data", "cost_diversion.xlsx"),
          sheet = "commercial"
          ) %>% 
  gt() %>% 
  cols_label(
    flight = "Type of flight",
    min = "Cost per diverted flight",
    max = "Cost per diverted flight",
    recommended = "Recommended value...",
    pax_compensation = "...of which passenger compensation"
  ) %>%
  fmt_currency(
    currency = "EUR",
    columns = min:pax_compensation,
    decimals = 0
  ) %>% 
  cols_merge_range(min, max, sep = "--", autohide = TRUE) %>% 
   tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
     flight ~ pct(30),
     min:pax_compensation ~ pct(23)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
  footnote = "Passenger compensation refers solely to the passenger compensation claimed under Regulation (EC) 261/2004",
locations = cells_column_labels(
  columns = pax_compensation
)
) %>% 
  tab_footnote(
    footnote = "LCC stands for Low-Cost Carrier",
    locations = cells_body(columns = flight, rows = 2),
    placement = "right"
  ) %>% 
  tab_source_note(
    source = md("*Source: Values estimated in consultation with Airlines*")
  )
```

The values in @tbl-diversion-comm are presented for three different flight types,
which also correspond to different aircraft types, typically flown on such routes:

* **Regional flights** refer to short-haul flights (< 1,500 km) and that are 
typically operated by small aircraft (80-120 seats), such as regional jets and turboprop

* **Continental flights** refer to medium-haul flights (1,500 – 3,000 km) and that
are typically operated by narrow-body aircraft (130 – 189 seats). The continental
flights are split between those operated by Mainline carriers and Low-Cost carriers 
(LCC), in order to better reflect the respective cost differences

* **Intercontinental flights** refer to long-haul flights (> 3,000 km) and that 
are typically operated by wide-body aircraft (200 – 500 seats). These are most 
often flights departing outside of ECAC area and the focus of the analysis is on
flights that are diverted within the ECAC territory for simplicity

Furthermore, three categories of values are provided:

* **Cost per diverted flight**, in the first column, represents a cost interval per
diverted flight. Within this cost interval the following assumptions are made:

  + *For the lower end of the interval:* the flight is diverted to an airport 
  where the Airline has assistance contracts in place; the additional flight 
  time to divert is short; there are few passengers in connection on board and 
  there is no need for passenger compensation, only a meal voucher is provided
  
  + *For the upper end of the interval:* the flight is diverted to an airport
  where the Airline has no assistance contracts in place; additional flight time
  is longer and more expensive; there are several passengers in connection on board;
  the delay caused is long, triggering [Regulation (EC) 261/2004](https://eur-lex.europa.eu/eli/reg/2004/261/oj/eng);
  both a meal voucher   and a bus or taxi transfer, or a night at a hotel is
  provided to the passengers;  the diversion causes a delay for the following 
  flight, requiring passenger care to be provided to the passengers on the next 
  flight; and a ferry flight is required in order to reposition the aircraft for further operations
  
* **Recommended value** presents an estimated most likely cost of diversion per
type of flight, weighted by the likelihood of occurrence of the drivers impacting
the cost per flight. This value includes the cost of passenger compensation as per
[Regulation (EC) 261/2004](https://eur-lex.europa.eu/eli/reg/2004/261/oj/eng),
and aims at guiding the user towards a most likely cost per flight within the 
interval provided in column “Cost per diverted flight”

* **Of which passenger compensation** presents the amount corresponding to
passenger compensation within the value provided in the column “Recommended value”,
per [Regulation (EC) 261/2004](https://eur-lex.europa.eu/eli/reg/2004/261/oj/eng). 
This metric is separated as this is a differentiating factor between the different
uses of the cost of diversion (i.e. those due to Airline internal reasons, where 
the regulation is applicable, and those due to reasons external to the Airline,
where the regulation is not applicable)

In 2024, out of the total number of flights (9.4 million) with a destination in
the EUROCONTROL Network Manager area, over 30 thousand flights (0.32%) landed at
an airport other than the one initially planned.

The **cost of diversion for business aviation**, as presented below, has not been
recalculated due to lack of available information. Therefore, the value reflects
the initial estimation from 2012 adjusted to inflation.

```{r}
#| label: tbl-diversion-bus
#| tbl-cap: Estimated cost of diversion for business aviation

read_xlsx(here::here("data", "cost_diversion.xlsx"),
          sheet = "business"
          ) %>% 
  gt() %>% 
  cols_label(
    type = "Type of flight",
    cost = "Cost of diverted flight"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = cost,
    decimals = 0
  ) %>% 
   tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
    everything() ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
  footnote = "Monetary values were adjusted from 2012 to 2024 prices according to inflation",
locations = cells_column_labels(
  columns = cost
)
) %>% 
  tab_source_note(
    source = md("*Source: Data supplied by the airline members of the SESAR CBA team (2015)*")
  )
```

:::{.callout-note}
The estimated cost for business aviation assumes that for each diverted flight 
there is one additional positioning flight.
:::

## Other possible values

**Cost of diversion due to curfew** represents the estimated cost per flight when
a diversion occurs in order to avoid a curfew at the airport of destination.

```{r}
#| label: tbl-diversion-curfew
#| tbl-cap: Estimated average cost of a diversion due to curfew for scheduled aviation

read_xlsx(here::here("data", "cost_diversion.xlsx"),
          sheet = "curfew"
          ) %>% 
  gt() %>% 
  cols_label(
    flight = "Type of flight",
    min = "Cost per diverted flight",
    max = "Cost per diverted flight",
    recommended = "Recommended value...",
    pax_compensation = "...of which passenger compensation"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = min:pax_compensation,
    decimals = 0
  ) %>% 
  cols_merge_range(min, max, sep = "--", autohide = TRUE) %>% 
   tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
     flight ~ pct(30),
     min:pax_compensation ~ pct(23)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
  footnote = "Passenger compensation refers solely to the passenger compensation claimed under Regulation (EC) 261/2004",
locations = cells_column_labels(
  columns = pax_compensation
)
) %>% 
  tab_footnote(
    footnote = "LCC stands for Low-Cost Carrier",
    locations = cells_body(columns = flight, rows = 2),
    placement = "right"
  ) %>% 
  tab_source_note(
    source = md("*Source: Values estimated in consultation with Airlines*")
  )
```

The logic behind the estimation of ‘cost of diversion due to curfew’ is identical 
to the one described for the cost of diversion above. However, diversions
due to a curfew have some differences from other en-route diversions in that 
they presume that the diversion happens late in the evening, which entails a set
of costs not necessarily faced under other causes of diversion. Therefore, the
main difference in cost estimation is the assumption that there would be a need 
for bus/taxi transfer and/or hotel accommodation for the passengers; there will 
be reactionary delay the following day due to the positioning of the aircraft at
an airport other than that initially planned, leading to costs for passenger care
for passengers on the next day’s flight; and that the passengers would claim 
compensation in any scenario (i.e. in the lower end of the cost interval and in 
the higher end), provided that the diversion falls under [Regulation (EC) 261/2004](https://eur-lex.europa.eu/eli/reg/2004/261/oj/eng).