# Cost of diversion {#sec-cost-of-diversions}

## EUROCONTROL recommended values

```{r}
library(readxl)
library(dplyr)
library(gt)
```

@tbl-diversion-comm and @tbl-diversion-bus present the estimated cost of diversion of a flight to an airport
other than the one initially planned. The values are split between commercial and business 
aviation, and, where available, represent a range of values as estimated by the 
airline members consulted.

```{r}
#| label: tbl-diversion-comm
#| tbl-cap: Estimated cost of diversion for commercial aviation

read_xlsx(here::here("data", "cost_diversion.xlsx"),
          sheet = "commercial"
          ) %>% 
  gt() %>% 
  cols_label(
    type = "Type of flight",
    min = "Cost of diverted flight",
    max = "Cost of diverted flight"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = min:max,
    decimals = 0
  ) %>% 
  cols_merge_range(min, max, sep = "--", autohide = TRUE) %>% 
   tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
    everything() ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
  footnote = "Monetary values were adjusted from 2006 to 2022 prices according to inflation",
locations = cells_column_labels(
  columns = min
)
) %>% 
  tab_source_note(
    source = md("*Source: Data supplied by the airline members of the SESAR evaluation team, derived from an analysis of 2006 ECAC data*")
  )
```

:::{.callout-note}
The penalties associated with the late delivery of cargo are not considered in the estimation, as this type of data is not yet readily available.
:::

```{r}
#| label: tbl-diversion-bus
#| tbl-cap: Estimated cost of diversion for business aviation

read_xlsx(here::here("data", "cost_diversion.xlsx"),
          sheet = "business"
          ) %>% 
  gt() %>% 
  cols_label(
    type = "Type of flight",
    cost = "Cost of diverted flight"
  ) %>% 
  fmt_currency(
    currency = "EUR",
    columns = cost,
    decimals = 0
  ) %>% 
   tab_options(
    table.width = pct(100)
    ) %>% 
   cols_width(
    everything() ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
  footnote = "Monetary values were adjusted from 2012 to 2022 prices according to inflation",
locations = cells_column_labels(
  columns = cost
)
) %>% 
  tab_source_note(
    source = md("*Source: Data supplied by the airline members of the SESAR CBA team (2015)*")
  )
```

:::{.callout-note}
The estimated cost for business aviation assumes that for each diverted flight there is one additional positioning flight.
:::

In 2022, out of the total number of flights (9.3 million) with a destination in EUROCONTROL Network Manager area, 28,738 flights (0.3%) landed at an airport other than the one initially planned.

