# Distance flown by charging zone {#sec-distance-flown-by-charging-zone}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

@tbl-distance-flown shows the million kilometres flown by charging zone.

```{r}
#| label: tbl-distance-flown
#| tbl-cap: Distance flown by charging zone in million km

read_xlsx(here::here("data", "distance_by_charging_zone.xlsx")) %>% 
  gt() %>% 
  cols_label(
    State = "State",
    y2014 = "2014",
    y2015 = "2015",
    y2016 = "2016",
    y2017 = "2017",
    y2018 = "2018",
    y2019 = "2019",
    y2020 = "2020",
    y2021 = "2021",
    y2022 = "2022"
  ) %>% 
  fmt_number(
    columns = y2014:y2022,
    decimals = 0
  ) %>% 
  cols_align(
    align = "right",
    columns = y2014:y2022
  ) %>% 
   tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = State:y2022,
      rows = 39
    )
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = State:y2022,
      rows = 43
    )
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>%  
   cols_width(
    y2014:y2022 ~ pct(8),
    State ~ pct(20),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source = md("*Source: [EUROCONTROL PRU (n.d), ATM Cost Effectiveness (ACE) Reports](https://ansperformance.eu/publications/prc/ace/)*")
  ) %>% 
  tab_footnote(
    footnote = "Estonia integrated as of April 2017",
    locations = cells_body(
      columns = State,
      rows = 40
    )
    ) %>%  
  tab_footnote(
    footnote = "Ukraine integrated as of November 2021",
    locations = cells_body(
      columns = State,
      rows = 41
    )
    ) 
```

For the most updated information, please refer to the [Aviation Intelligence En-Route Dashboard](https://www.eurocontrol.int/ServiceUnits/Dashboard/EnRouteMainDashboard.html).

The Report on the Operation of the Route Charges System is published by the CRCO 
on an annual basis and provides data on traffic volumes and ATM costs for the 
States for which the CRCO collects en-route and terminal charges.

@tbl-distance-flown sets out the number of millions of kilometres recorded in 
the airspace of the Contracting States from 2014 to 2022 for the calculation of
route charges (great circle distance after deduction of 20 km for departing and 
arriving flights).

Information on the various charges levied by the CRCO, in particular the charge 
calculation methods, the basic billing documents, the methods of payment and the
submission of claims is described in  [customer guide to charges](https://www.eurocontrol.int/publication/customer-guide-route-charges).

## When to use the input?

This input is suitable for use in assessments aiming at estimating the airspace 
users’ operating costs, namely when it comes to the route charges.

## Related inputs

@sec-if-average-flight-distance-and-flight-duration [IFR average flight distance and duration](#sec-if-average-flight-distance-and-flight-duration)

@sec-en-route-ans-costs [En-route ANS costs](#sec-en-route-ans-costs)

@sec-route-charge-share-per-market-segment [Route charge share per market segment](sec-route-charge-share-per-market-segment)

