```{r}
library(readxl)
library(dplyr)
library(gt)
```

# Fleet CNS capability {#sec-fleet-cns-capability}

## EUROCONTROL recommended values

The subsections below provide statistics on flights and aircraft with certain 
Communication, Navigation and Surveillance (CNS) equipment and capabilities.

The [EUROCONTROL Aircraft communication, navigation & surveillance (CNS) dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard)
provides information for monitoring fleet capabilities and preparing performance-based
navigation (PBN) deployment plans. It does so by analysing CNS and PBN information
contained in ICAO flight plans and generates reports on aircraft or flight 
characteristics. Note that the capabilities of aircraft which do not submit flight
plans are not covered in the dashboard figures. The missing information is to a 
large extent for general aviation (GA) flights. 

The dashboard provides statistics on equipment and capabilities such as:

* Communication: FMC WPR ACARS, HF RTF; CPDLC FANS 1/A HFDL; etc.

* Navigation: RNAV 5, RNAV 1, RNP 1, RNP APCH (including LPV capability), GBAS, etc.

* Surveillance: ADS-B, ADS-C, Mode S transponder, etc.

Different periods of time, airports, airlines, or aircraft types 
(depending on the user profile) can be analysed. 

To access the dashboard, you first need to register on the OneSky Online portal
using the link in the source above.

### Navigation (NAV) flight capabilities

The values presented in the tables hereafter are weighted by flight and represent
the values for 2022. All values come from [EUROCONTROL Aircraft communication, navigation & surveillance (CNS) dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard).
Please refer to this source for more detailed information and further data.

```{r}
#| label: tbl-nav-flight
#| tbl-cap: NAV flight capabilities as a % of total flights in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "nav_total") %>% 
  gt() %>% 
  cols_label(
    capability = "Capability",
    flights = "Number of flights",
    share_flights = "As a % of all flights"
  ) %>% 
  fmt_number(
    columns = flights,
    decimals = 0
  ) %>% 
  fmt_percent(
    columns = share_flights,
    decimals = 2
  ) %>% 
  cols_align(
    align = "right",
    columns = flights
  ) %>% 
  cols_align(
    align = "center",
    columns = share_flights
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(33),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "The number of flights is the number of arrivals or departures, or arrivals and departures in line with the selection criteria in the CNS dashboard",
    locations = cells_column_labels(
      columns = flights
    )
  ) 
```

```{r}
#| label: tbl-nav-flight_top10
#| tbl-cap: NAV flight capabilities as a % at the top 10 European Airports in Q4 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "nav_top10") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    gbas = "A - GBAS landing system",
    lpv = "B - LPV (APV with SBAS)",
    rnav = "D - RNAV 1",
    rnp = "O - Basic RNP 1",
    rnp_apch = "S - RNP APCH"
  ) %>% 
  fmt_percent(
    columns = gbas:rnp_apch,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = gbas:rnp_apch
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    airport ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "The values for GBAS are corrected values. They exclude DHC8 equipment only compatible with a GBAS precursor system",
    locations = cells_column_labels(
      columns = gbas
    )
  ) 
```

```{r}
#| label: tbl-nav-flight_month
#| tbl-cap: NAV aircraft capability at the top-10 European airports over a one-month period in Q4 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "nav_month") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    gbas = "A - GBAS landing system",
    lpv = "B - LPV (APV with SBAS)",
    rnav = "D - RNAV 1",
    rnp = "O - Basic RNP 1",
    rnp_apch = "S - RNP APCH"
  ) %>% 
  fmt_percent(
    columns = gbas:rnp_apch,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = gbas:rnp_apch
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    airport ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

```{r}
#| label: tbl-nav-flight_aircraft
#| tbl-cap: NAV flight capability by main aircraft model, December 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "nav_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft",
    gbas = "A - GBAS landing system",
    lpv = "B - LPV (APV with SBAS)",
    rnav = "D - RNAV 1",
    rnp = "O - Basic RNP 1",
    rnp_apch = "S - RNP APCH"
  ) %>% 
  fmt_percent(
    columns = gbas:rnp_apch,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = gbas:rnp_apch
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(15),
    aircraft ~ pct(25)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) 
```

### Communication (COM) flight capabilities

The values presented in the tables hereafter are weighted by flight. Values are 
based on flights arriving at ECAC Airports and the data is originated from
[EUROCONTROL Aircraft communication, navigation & surveillance (CNS) dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard). 
Please refer to this source for more detailed information and further data.

```{r}
#| label: tbl-com-flight
#| tbl-cap: COM flight capabilities as a % of total flights in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "com_total") %>% 
  gt() %>% 
  cols_label(
    capability = "Capability",
    flights = "Number of flights",
    share_flights = "As a % of all flights"
  ) %>% 
  tab_row_group(
    label = "Voice",
    rows = 7:13
  ) %>% 
  tab_row_group(
    label = "Datalink",
    rows = 1:6
  ) %>% 
  tab_options(
    row_group.background.color = "#F5F7F9",
    row_group.font.weight = "bold"
  ) %>% 
  fmt_number(
    columns = flights,
    decimals = 0
  ) %>% 
  fmt_percent(
    columns = share_flights,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = share_flights
  ) %>% 
  cols_align(
    align = "right",
    columns = flights
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(30),
    capability ~ pct(40)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

```{r}
#| label: tbl-com-flight_top10
#| tbl-cap: COM datalink flight capabilities as a % at the Pilot Common Project airports in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "com_top10") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    fis_acars = "E2 - D-FIS ACARS",
    pdc_acars = "E3 - PDC ACARS",
    cpdlc_atn_cdl = "J1 - CPDLC ATN VDL Mode 2",
    cpdlc_fans_vld = "J4 - CPDLC FANS 1/A VDL Mode 2",
    inmarsat = "J5 - CPDLC FANS 1/A SATCOM (INMARSAT)",
    iridium = "J7 - CPDLC FANS 1/A SATCOM (IRIDIUM)"
  ) %>% 
  fmt_percent(
    columns = fis_acars:iridium,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = fis_acars:iridium
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(8),
    airport ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) 
```

```{r}
#| label: tbl-com-voice-top10
#| tbl-cap: COM voice flight capabilities as a % at the Pilot Common Project Airports in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "com_voice_top10") %>% 
  gt() %>% 
  cols_label(
    airport = "Airport",
    hf_rtf = "H – HF RTF",
    inmarsat = "M1 – ATC RTF SATCOM (INMARSAT)",
    mtsat = "M2 – ATC RTF (MTSAT)",
    iridium = "M3 – ATC RTF (IRIDIUM)",
    uhf_rtf = "U – UHF RTF",
    vhf_rtf = "V – VHF RTF",
    vhf_spacing = "Y – VHF with 8.33 kHz channel spacing"
  ) %>% 
  fmt_percent(
    columns = hf_rtf:vhf_spacing,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = hf_rtf:vhf_spacing
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(9),
    airport ~ pct(37)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) 
```
 
```{r}
#| label: tbl-com-aircraft
#| tbl-cap: COM datalink flight capabilities for main aircraft models in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "com_datalink_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft",
    fis_acars = "E2 - D-FIS ACARS",
    pdc_acars = "E3 - PDC ACARS",
    cpdlc_atn_cdl = "J1 - CPDLC ATN VDL Mode 2",
    cpdlc_fans_vld = "J4 - CPDLC FANS 1/A VDL Mode 2",
    inmarsat = "J5 - CPDLC FANS 1/A SATCOM (INMARSAT)",
    iridium = "J7 - CPDLC FANS 1/A SATCOM (IRIDIUM)"
  ) %>% 
  fmt_percent(
    columns = fis_acars:iridium,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = fis_acars:iridium
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(14)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) 
```
 
```{r}
#| label: tbl-com-voice-aircraft
#| tbl-cap: COM voice flight capabilities for main aircraft models in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "com_voice_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft type",
    hf_rtf = "H – HF RTF",
    inmarsat = "M1 – ATC RTF SATCOM (INMARSAT)",
    mtsat = "M2 – ATC RTF (MTSAT)",
    iridium = "M3 – ATC RTF (IRIDIUM)",
    uhf_rtf = "U – UHF RTF",
    vhf_rtf = "V – VHF RTF",
    vhf_spacing = "Y – VHF with 8.33 kHz channel spacing"
  ) %>% 
  fmt_percent(
    columns = hf_rtf:vhf_spacing,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = hf_rtf:vhf_spacing
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(9),
    aircraft ~ pct(37)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

### Surveillance (SUR) flight capabilities

The values presented in the tables hereafter refer to the level of flight capabilities
in terms of surveillance equipment. All values are weighted by flight and represent
the situation for 2022. The data used is originated from
[EUROCONTROL Aircraft communication, navigation & surveillance (CNS) dashboard](https://www.eurocontrol.int/dashboard/communication-navigation-and-surveillance-dashboard). 
Please refer to this source for more detailed information and further data.

```{r}
#| label: tbl-sur-flight
#| tbl-cap: SUR flight capabilities as a % of total flights in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "sur_total") %>% 
  gt() %>% 
  cols_label(
    capability = "Capability",
    flights = "Number of flights",
    share_flights = "As a % of all flights"
  ) %>% 
  tab_row_group(
    label = "ADS-B",
    rows = 10:15
  ) %>% 
  tab_row_group(
    label = "SSR Mode S",
    rows = 1:9
  ) %>% 
  tab_options(
    row_group.background.color = "#F5F7F9",
    row_group.font.weight = "bold"
  ) %>% 
  fmt_number(
    columns = flights,
    decimals = 0
  ) %>% 
  fmt_percent(
    columns = share_flights,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = share_flights
  ) %>% 
  cols_align(
    align = "right",
    columns = flights
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    capability ~ pct(70),
    flights ~ pct(15),
    share_flights ~ pct(15)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
    footnote = "Figures in this value are per flight from departure to destination, while values per airport are per flight but also include any flight departing from or arriving at that airport",
    locations = cells_column_labels(
      columns = flights
    )
  )
```

```{r}
#| label: tbl-ssr-top10
#| tbl-cap: SSR/Mode S declared capabilities in flight plans at the top 10 European airports in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "ssr_top10") %>% 
  gt() %>% 
  tab_spanner(
    "Capability",
    columns = L:A
  ) %>% 
  fmt_percent(
    columns = L:A,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = L:A
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    L:A ~ pct(9),
    Airport ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

```{r}
#| label: tbl-adsb-top10
#| tbl-cap: ADS-B declared capabilities in flight plans at the top 10 European airports in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "adsb_top10") %>% 
  gt() %>% 
  tab_spanner(
    "Capability",
    columns = B1:V2
  ) %>% 
  fmt_percent(
    columns = B1:V2,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = B1:V2
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    B1:V2 ~ pct(14),
    Airport ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) 
```

```{r}
#| label: tbl-ssr-aircraft
#| tbl-cap: SSR/Mode S declared capabilities in flight plans for main aircraft models in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "ssr_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft type"
  ) %>% 
  tab_spanner(
    "Capability",
    columns = L:A
  ) %>% 
  fmt_percent(
    columns = L:A,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = L:A
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    L:A ~ pct(9),
    aircraft ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) 
```

```{r}
#| label: tbl-adsb-aircraft
#| tbl-cap: ADS-B declared capabilities in flight plans for main aircraft models in 2022

read_xlsx(here::here("data", "cns_capability.xlsx"),
          sheet = "adsb_aircraft") %>% 
  gt() %>% 
  cols_label(
    aircraft = "Aircraft type"
  ) %>% 
  tab_spanner(
    "Capability",
    columns = B1:V2
  ) %>% 
  fmt_percent(
    columns = B1:V2,
    decimals = 1
  ) %>%
  cols_align(
    align = "center",
    columns = B1:V2
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    B1:V2 ~ pct(14),
    aircraft ~ pct(19)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

## Comment

The numbers of flights and aircraft provided by the CNS dashboard are derived from 
flight plans submitted to the EUROCONTROL Network Manager. Consequently, the 
statistics do not include the capability of aircraft flying in uncontrolled 
airspace or under VFR and thus do not submit flight plans to the NM. On-board 
capability and equipment data made available via the CNS dashboard are those 
declared in ICAO FPLs by operators. The information is therefore only as reliable
as declared. For detailed analysis, additional local assessment is recommended.

## Related inputs

@sec-number-of-ifr-flights [Number of IFR flights](#sec-number-of-ifr-flights)

 @sec-cns-infrastructure [CNS infrastructure](#sec-cns-infrastructure)

@sec-pbn-and-precision-approach-procedures [PBN and precision approach procedure](#sec-pbn-and-precision-approach-procedures)

