# Fleet size {#sec-fleet-size}

```{r}
library(readxl)
library(dplyr)
library(gt)
library(tibble)
```


This input presents the number of aircraft, per type, operating in Europe.

@tbl-fleet-size shows the top 30 civil aircraft operating in 2022 in the airspace controlled
by the EUROCONTROL Network Manager, by aircraft type and number of aircraft.

```{r}
#| label: tbl-fleet-size
#| tbl-cap: Top 30 civil aircraft operating in NM area

read_xlsx(here::here("data", "fleet_size.xlsx"),
          sheet = "top30") %>% 
  gt() %>% 
  cols_label(
    type = "Aircraft type",
    number = "Number of aircraft",
    flights_2022 = "Number of flights",
    proportion_aircraft = "Proportion of aircraft",
    cum_proportion_aircraft = "Cumulative",
    proportion_flights = "Proportion of flights",
    cum_proportion_flights = "Cumulative"
  ) %>% 
  fmt_number(
    columns = number:flights_2022,
    decimals = 0
  ) %>% 
  fmt_percent(
    columns = proportion_aircraft:cum_proportion_flights,
    decimals = 2
  ) %>% 
  cols_align(
    align = "center",
    columns = proportion_aircraft:cum_proportion_flights
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(14),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 32
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: EUROCONTROL Network Manager flight plans, 2022*")
  )
```

The 22,948 civil aircraft in @tbl-fleet-size include 500 unique aircraft types. 
The top 30 aircraft types listed above represent approximately 62% of the total 
fleet, while 69% of flights are operated by the 300 most used aircraft types.

@tbl-military-aircraft shows the number of military aircraft by category in 2023 vs 2019 
in Europe, USA, and Commonwealth of Independent States (CIS) countries (Armenia,
Azerbaijan, Belarus, Kazakhstan, Kyrgyzstan, Moldova, Russia, Tajikistan and Uzbekistan).[@flightglobal2023]

```{r}
#| label: tbl-military-aircraft
#| tbl-cap: Military aircraft in 2023 vs 2019

read_xlsx(here::here("data", "fleet_size.xlsx"),
          sheet = "military") %>% 
  gt() %>% 
  cols_label(
    aircraft_type = "Aircraft type",
    europe2019 = "Europe 2019",
    europe2023 = "Europe 2023",
    usa2019 = "USA 2019",
    usa2023 = "USA 2023",
    cis2019 = "CIS 2019",
    cis2023 = "CIS 2023"
  ) %>% 
  fmt_number(
    columns = europe2019:cis2023,
    decimals = 0,
    rows = 1:7
  ) %>% 
  fmt_percent(
    columns = europe2019:cis2023,
    decimals = 2,
    rows = 8
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    aircraft_type ~ pct(22),
    europe2019:cis2023  ~ pct(13)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 7
    )
  ) %>% 
  tab_style(
    style = list(
      cell_text(
        style = "italic"
      )
    ),
    locations = cells_body(
      columns = everything(),
      rows = 8
    )
  ) %>% 
  tab_footnote(
    footnote = "Special Mission Aircraft are those platforms specifically developed to undertake an over-battlefield role by utilization of advanced onboard equipment or specialized trait",
    locations = cells_body(
      columns = aircraft_type,
      rows = 2
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [FlightGlobal (2023) 2023 World Air Forces directory](https://www.flightglobal.com/reports/2023-world-air-forces-directory/151088.article)*")
  )
```

@tbl-ifr-helicopter shows the fleet size of IFR helicopters in Europe for years 2011/2012.

```{r}
#| label: tbl-ifr-helicopter
#| tbl-cap: IFR helicopter fleet

tribble(
 ~Region, ~Units,
  "Europe and Eastern Europe",  "2,208",
  "CIS countries",              "1,312",
  "Total",                      "3,520"
) %>% 
  gt() %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 3
    )
  ) %>% 
  tab_options(
    table.width = pct(60)
   ) %>% 
   cols_width(
    everything() ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
   tab_source_note (
    source = md("*Source:  FlightGlobal HELICAS 2011/2012 data analysed by the [European Helicopter Association](https://ext.eurocontrol.int/ground_navigation_infrastructure/homepage/welcome)*")
  )
```

@tbl-airframes estimates the number of airframes operating under Visual Flight Rules (VFR) in ECAC region.

```{r}
#| label: tbl-airframes
#| tbl-cap: Number of VFR airframes in ECAC

tribble(
 ~Aircraft, ~Units,
  "Aeroplanes",  "20,329",
  "Helicopters", "3,532",
  "Gliders", "18,555",
  "Balloons", "6,623",
  "Total", "49,039"
) %>% 
  gt() %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = everything(),
      rows = 5
    )
  ) %>% 
  tab_options(
    table.width = pct(60)
   ) %>% 
   cols_width(
    everything() ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
   tab_source_note (
    source = md("*Source: [International Registry of Civil Aircraft (IRCA), 2014](https://www.icao.int/secretariat/legal/Pages/Intl_registry.aspx) and [Aircraft Owners and Pilot Association (AOPA) (2015)](https://www.aopa.org/)*")
  )
```

## Other possible sources - forward-looking

The **Airbus Global Market Forecast for 2022-2041** presents a forward-looking view
of the evolution of the air transport sector, accounting for factors such as 
demographic and economic growth, tourism trends, oil prices, the development of 
new and existing routes, and ultimately highlighting demand for aircraft covering 
the spectrum of sizes from 100 seats to the very largest aircraft of over 500 seats.

![Airbus commercial aircraft demand 2022-2041 @airbusGMF2022](../figures/airbus_demand.svg){#fig-airbus-demand}

As shown in @fig-airbus-demand, for the years 2022-2041, Airbus forecasts a 
demand for 39,490 new or replacement commercial aircraft worldwide, of which 
8,140 units (21%) are for Europe/CIS regions. 
Asia-Pacific, China and the US are driving the growth, while single-aisle
aircraft dominate the demand (>80%).

@fig-boeing-demand shows the Boeing forecast demand for new aircraft by category 
and world regions with a focus on single-aisle aircraft which leads the market growth.

![Boeing commercial aircraft demand 2022-2041 @boeingcmf2022](../figures/boeing_demand.svg){#fig-boeing-demand}

For the same period (2022-2041), **Boeing forecasts** a global demand for 40,170
new or replacement commercial aircraft of which 8,550 units (21%) are for Europe.
Europe, Asia-Pacific, US and China are the drivers for growth owing to the demand
for single-aisle aircraft which represents between 69% (in US) and 82% (in Europe)
of the total fleet forecast.

## Related inputs


@sec-number-of-ifr-flights [Number of IFR flights](#sec-number-of-ifr-flights)

@sec-ifr-flight-information-per-market-segment [IFR flight information per market segment](#sec-ifr-flight-information-per-market-segment)

@sec-fleet-age [Fleet age](#sec-fleet-age)

## References
