---
editor: 
  markdown: 
    wrap: 72
---

# Passenger value of time {#sec-passenger-value-of-time}
```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

```{r}
#| label: tbl-pas-time-value-eu
#| tbl-cap: Estimated value of travel time - EU25

read_xlsx(here::here("data", "passenger_time_value.xlsx"),
          sheet = "eu25") %>% 
  gt() %>% 
  cols_label(
    purpose = "Travel purpose",
    value = "Time value (€/hour)"
  ) %>% 
  fmt_currency(
   columns = c(value),
   currency = "EUR",
   decimals = 2
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_footnote(
  footnote = "Values adjusted from 2002 to 2024 prices based on inflation",
locations = cells_column_labels(
  columns = value
)
) %>% 
  tab_source_note(
    source_note = md("*Source: [European Commission (2006), HEATCO, Developing Harmonised European Approaches for Transport Costing and Project Assessment – Deliverable 5 Proposal for Harmonised Guidelines](https://trimis.ec.europa.eu/sites/default/files/project/documents/20130122_113653_88902_HEATCO_D5_summary.pdf)*")
  )
```

```{r}
#| label: tbl-pas-time-value-fr
#| tbl-cap: Estimated value of travel time - France
read_xlsx(here::here("data", "passenger_time_value.xlsx"),
          sheet = "france") %>% 
  gt() %>% 
  cols_label(
    purpose = "Travel purpose",
    value = "Time value (€/hour)"
  ) %>% 
  fmt_currency(
   columns = c(value),
   currency = "EUR",
   decimals = 2
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [French Ministry of Ecological Transition (2019), Recommended values for calculating average long-distance travel](https://www.ecologie.gouv.fr/sites/default/files/V.3.pdf)*")
  ) %>%
  tab_footnote(
  footnote = "Values adjusted from 2015 to 2024 prices based on inflation",
locations = cells_column_labels(
  columns = value
)
)
  
```

```{r}
#| label: tbl-pas-time-value-uk
#| tbl-cap: Estimated value of travel time - United Kingdom
read_xlsx(here::here("data", "passenger_time_value.xlsx"),
          sheet = "uk") %>% 
  gt() %>% 
  cols_label(
    purpose = "Travel purpose",
    value = "Time value (€/hour)"
  ) %>% 
  fmt_currency(
   columns = c(value),
   currency = "EUR",
   decimals = 2
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [Airports Commission (2014), Values of time used to estimate passenger delay costs in the UK airport system](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/372606/AC08a_tagged.pdf)*")
  ) %>% 
  tab_footnote(
  footnote = "Values adjusted from 2014 to 2024 prices based on inflation and using the exchange rate in general parameters",
locations = cells_column_labels(
  columns = value
)
) 
  
```

This input provides an estimation of the average value of passenger time spent 
travelling, which might alternatively be spent on other activities (e.g. working
or leisure). It is essentially the opportunity cost which corresponds to the monetary
value associated with a passenger during a journey. It shows how much a passenger 
would be willing to pay in order to save time during a journey (e.g. by travelling 
on a quicker service or using a faster transport mode), or how much ‘compensation’
they would accept, directly or indirectly, for time lost. @tbl-pas-time-value-eu,
@tbl-pas-time-value-fr and @tbl-pas-time-value-uk present a collection of different 
values that can be used for this purpose. It is to be noted that, in this section, 
the value of time is not cited as a function of delay duration. This is an important 
consideration when using the value, since the longer the delay duration, the higher the value.

## Comment

When looking into the values in the table, a few points regarding the sources 
should be taken into consideration:

* The source used for the UK values sets out a methodology for analysis which has
  been undertaken to estimate benefits from reduced delay time to airlines and 
  passengers from changes in aviation capacity constraints in the UK for 11 airports.
  
* The values for France rely on a working paper on recommended values for 
  calculating the components of a socio-economic net present value, which include 
  travel time. The assessment therefore covers social, environmental, and economic 
  effects.

* Regarding the numbers for EU25, they remain a reference if a European value is
  sought. The objective of the study from which they were derived is to propose 
  harmonised guidelines for project assessment for transnational projects in Europe.
  It provides monetary estimates for the values of time saved for an employer’s
  business and for passenger non-work trips (e.g. commuting, shopping and leisure).

## Other possible values

Some additional values, which, although not constituting the perceived key inputs, 
may be useful for specific purposes of the user of these inputs. These values are 
presented below.

***Value of time of a business aviation passenger***

**€153 per hour.** This value was provided to EUROCONTROL by airline members of 
the SESAR CBA team in 2012 and is, therefore, adjusted from 2011 prices. 

***Value of passenger time in the US for high-speed rail passengers***

```{r}
#| label: tbl-pas-time-value-us
#| tbl-cap: Passenger value of time for air and high-speed rail travel
read_xlsx(here::here("data", "passenger_time_value.xlsx"),
          sheet = "additional") %>% 
  gt() %>% 
  cols_label(
    purpose = "Travel purpose",
    value = "Time value (€/hour)"
  ) %>% 
  fmt_currency(
   columns = c(value),
   currency = "EUR",
   decimals = 2
  ) %>% 
  
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ px(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: US Department of Transportation (2016), Revised Departmental Guidance on Valuation of Travel Time in Economic Analysis*")
  ) %>%
  tab_footnote(
  footnote = "Values adjusted from 2015 to 2024 prices based on inflation and using the exchange rate in general parameters",
locations = cells_column_labels(
  columns = value
)
) 
  
```

Based on US Department of Transport (DOT) guidance on passenger value of time 
for air and high-speed rail travel by purpose of trip. The numbers are adjusted 
to inflation from 2015 to 2024 prices.

## Further reading

Below are listed some additional sources recommended for consultation:

* *International Transport Forum (ITF), “What is the Value of Saving Travel Time?” Feb. 2019[@itf:2019]*

* *Economic Development Research Group Inc., “Passenger Value of Time, Benefit-Cost Analysis and Airport Capital Investment Decisions,” 2015[@passtimevalue2015]*

* *University of Leeds, “Values of travel time in Europe: Review and meta-analysis,” 2016[@leedsuniversity2016]*

* *University of Leeds, “European Wide-Meta Analysis of Values of Travel Time,” May 2012[@leedsuniversity2012]*

## When to use the input?

This input is expected to become useful in any study that looks at the opportunity 
cost of the use of air transport, delays, cancellations, etc. It provides a
perspective on the impact that a change in the air transport can have on the passenger.

## References
