# PBN and precision approach procedures {#sec-pbn-and-precision-approach-procedures}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended values

The values hereafter present the proportion of runway ends in ECAC with published 
Performance-Based Navigation (PBN) instrument approach procedures fully implemented.
The data presented in the tables comes from [EUROCONTROL Performance Based Navigation Map Tool](https://www.eurocontrol.int/platform/performance-based-navigation-map-tool). 
Please refer to this source for any further information.

```{r}
#| label: tbl-pbn-deployment
#| tbl-cap: PBN approach deployment status, March 2023

read_xlsx(here::here("data", "pbn_procedures.xlsx"),
          sheet = "pbn_deployment") %>% 
  gt() %>% 
  cols_label(
    approach = "Approach type",
    runway_nb = "Runway ends covered",
    runway_share = "% of runway ends covered"
  ) %>% 
  fmt_percent(
    columns = runway_share,
    decimals = 0
  ) %>% 
  fmt_number(
    columns = runway_nb,
    decimals = 0
  ) %>% 
  cols_align(
    align = "center",
    columns = runway_share
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    runway_nb:runway_share ~ pct(25),
    approach ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>%
  tab_footnote(
    footnote = "’GLS’ does not include instrument approach procedures based on GBAS proprietary precursor systems",
    locations = cells_body(
      columns = approach,
      rows = 7
    )
    )  
```

```{r}
#| label: tbl-ils-deployment
#| tbl-cap: ILS Cat I, Cat II/III deployment status, March 2023

read_xlsx(here::here("data", "pbn_procedures.xlsx"),
          sheet = "ils_deployment") %>% 
  gt() %>% 
  cols_label(
    approach = "Approach type",
    runway_nb = "Runway ends covered",
    runway_share = "% of runway ends covered"
  ) %>% 
  fmt_percent(
    columns = runway_share,
    decimals = 0
  ) %>% 
  fmt_number(
    columns = runway_nb,
    decimals = 0
  ) %>% 
  cols_align(
    align = "center",
    columns = runway_share
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    runway_nb:runway_share ~ pct(25),
    approach ~ pct(50)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  )
```

The EUROCONTROL PBN Map Tool illustrates the deployment of PBN instrument 
approach procedures against objectives set in ICAO Assembly Resolution 37-11 
and the European Regulation on PBN (in particular Commission Implementing 
Regulation (EU) 2018/1048 of 18 July 2018). 

PBN approaches include instrument approach procedures compliant with the 
following navigation specifications of the PBN Manual (ICAO Doc 9613):

* RNP APCH

* RNP AR APCH

The PBN Map tool provides a list of the current and planned airports and runway 
ends covered by each type of approach. The tool gives information about:

* deployment progress since 2012, on the basis of actual publications

* future deployment trends based on the draft PBN Transition plans submitted to EUROCONTROL Network Manager for consultation

* the availability of PBN approaches with vertical guidance (APV) on all runway ends or on runway ends without precision landing (e.g. ILS, MLS or GBAS)

* the deployment status for ECAC, individual countries, Pilot Common Project (PCP) airports and individual airports

In March 2023 (on the basis of the AIRAC cycle, see comments), 1,676 runways ends
were equipped with instrument approach procedures.

## Comment

The PBN Map Tool is updated in accordance with publications for every AIRAC cycle^[For more details about the AIRAC cycles and the corresponding dates, please refer to the corresponding [EUROCONTROL page](https://www.eurocontrol.int/publication/aeronautical-information-regulation-and-control-airac-dates)].
It therefore provides up-to-date information on the current deployment status.


