# Taxi time {#sec-taxiing-time}

```{r}
library(readxl)
library(dplyr)
library(gt)
```

## EUROCONTROL recommended value

@tbl-taxi-time shows the mean duration, in minutes, of taxi times at airports, 
based on flights in the ECAC area.

```{r}
#| label: tbl-taxi-time
#| tbl-cap: Average taxi-in and taxi-out time (minutes)

read_xlsx(here::here("data", "taxi_time.xlsx"),
          sheet = "total") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
    all_in = "Taxi-in",
    large_in = "Taxi-in",
    medium_in = "Taxi-in",
    all_out = "Taxi-out",
    large_out = "Taxi-out",
    medium_out = "Taxi-out"
  ) %>% 
  tab_spanner(
    "All airports",
    columns = c(all_in, all_out)
  ) %>% 
  tab_spanner(
    "Large to very large airports",
    columns = c(large_in, large_out)
  ) %>% 
   tab_spanner(
    "Medium to small airports",
    columns = c(medium_in, medium_out)
  ) %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    all_in:medium_out ~ pct(15),
    year ~ pct(10)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  cols_align(
    align = "left",
    columns = year
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "year"
    )
  ) %>% 
  tab_footnote(
    footnote = "Large to very large airports: >150,000 movements",
    locations = cells_column_spanners(
      spanners = "Large to very large airports"
    )
  ) %>% 
  tab_footnote(
    footnote = "Medium to small airports: 149,999 to 15,000 movements",
    locations = cells_column_spanners(
      spanners = "Medium to small airports"
    )
  ) %>% 
  tab_source_note(
    source_note = md("*Source: Computed from actual data provided by the EUROCONTROL Central Office for Delay Analysis (CODA)*")
  )
```

Data in @tbl-taxi-time is based on actual data from CODA. The taxi-out time is 
defined as the time spent by a flight between its actual off-block time (AOBT) 
and actual take-off time (ATOT). The taxi-in time is defined as the time spent 
between its actual landing time (ALDT) and actual in-block time (AIBT). The 
taxi-in and taxi-out durations are calculated on the basis of data sent by airlines to CODA.

@tbl-taxi-time-additional shows the average additional taxi out time per 
departure in the top 30 airports in terms of movements.

```{r}
#| label: tbl-taxi-time-additional
#| tbl-cap: Average additional taxi-out time per departure

read_xlsx(here::here("data", "taxi_time.xlsx"),
          sheet = "additional") %>% 
  gt() %>% 
   tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(10),
    Year ~ pct(20)
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: [EUROCONTROL (2022), Performance Review Report (PRR) 2021](https://www.eurocontrol.int/publication/performance-review-report-prr-2021)*")
  )
```

Values in @tbl-taxi-time-additional are based on actual data from airports. 
The additional taxi-out time is a proxy for the average departure runway queuing
time on the outbound traffic flow during congested periods at airports. It is 
the difference between the actual taxi-out time of a flight and a statistically 
determined unimpeded taxi-out time^[The unimpeded taxi-out time is the taxi-out time in non-congested conditions at airports.] 
based on taxi-out times in periods of low traffic demand. There is one unimpeded 
time per stand/runway combination at each airport.

## Other possible sources

*EUROCONTROL CODA Taxi times - Summer and Winter reports* [@taxi:time] show, by 
airport, seasonal taxi time statistics for the IATA winter season and the IATA summer season:

* Taxi-in times

* Taxi-out times

* Taxi-out times by wake turbulence category

These taxi times are calculated using the airline reported actual off-block time,
actual take-off time, actual landing time and actual in-block time, providing 
the aviation community with seasonal benchmark values.

For additional granularity, taxi-out times by wake turbulence category are also 
offered for a number of airports.


## Related inputs

@sec-if-average-flight-distance-and-flight-duration [IFR average flight distance and flight duration](#sec-if-average-flight-distance-and-flight-duration)

@sec-ifr-flight-information-per-market-segment [IFR flight information per market segment](#sec-ifr-flight-information-per-market-segment)

## References
