# Turnaround time {#sec-turnaround-time}

```{r}
library(readxl)
library(dplyr)
library(gt)
```


## EUROCONTROL recommended values

Turnaround time represents the time taken for unloading and ground handling 
preparation for the return journey of an aircraft. This corresponds to the time 
during which the aircraft must remain parked at the gate, including air traffic 
flow management (ATFM) delay.

@tbl-turnaround-time presents the evolution of mean scheduled and mean actual 
turnaround time, in minutes, for medium and heavy aircraft. Please note that this
data was provided by EUROCONTROL Central Office for Delay Analysis (CODA) and 
can be accessed, together with a number of additional information, on the [MIRROR tool](https://www.eurocontrol.int/tool/mirror).[@coda:mirror]

```{r}
#| label: tbl-turnaround-time
#| tbl-cap: Scheduled vs. actual turnaround time in ECAC

read_xlsx(here::here("data", "turnaround_time.xlsx"),
          sheet = "time") %>% 
  gt() %>% 
  cols_label(
    year = "Year",
     scheduled1 = "Scheduled",
    actual1 = "Actual",
    scheduled2 = "Scheduled",
    actual2 = "Actual"
  ) %>% 
  cols_align(
    align = "right",
    columns = scheduled1:actual2
  ) %>% 
  cols_align(
    align = "left",
    columns = year
  ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = "year"
    )
  ) %>% 
   tab_spanner(
     "Heavy",
     columns = 2:3
   ) %>%
     tab_spanner(
     "Medium",
     columns = 4:5
   ) %>% 
   tab_footnote(
    footnote = "The heavy, medium and light aircraft categories relate to ICAO wake vortex categories based on the maximum certificated take-off mass: Heavy aircraft types of 136 000 kg (300 000 lb) or more; Medium aircraft types less than 136 000 kg (300 000 lb) and more than 7 000 kg (15 500 lb); Light aircraft types of 7 000 kg (15 500 lb) or less",
    locations = cells_column_spanners(
      spanners = everything()
    )
    ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(25)
  ) %>% 
   tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
tab_source_note(
  source_note = md("*Source: EUROCONTROL – Computed from data supplied by the airline members to CODA*")
)
  
```

## Description

The values presented in @tbl-turnaround-time are computed from data supplied by airlines to EUROCONTROL 
CODA. It includes the data on the following market segments: traditional scheduled, 
low-cost and charter.

The total ground time of an aircraft includes overnight stops, maintenance slots,
fire breaks, etc., so specific cut-off values are applied to obtain the turnaround
time. The turnaround cut-off time for wake turbulence category H (Heavy) is 180
minutes, and for M (Medium) 150 minutes.

The actual turnaround time represents the difference between the actual off-block
time (AOBT) of a departing flight and the actual in-block time (AIBT) of the same
aircraft on the previous inbound flight. The scheduled turnaround time is the 
difference between scheduled time of departure (STD) of the departing flight and 
the scheduled time of arrival (STA) of the same aircraft on the previous inbound flight.

## Other possible values

@tbl-turnaround-time-range presents, for 2019, an overview of turnaround time ranges for the
10th (Low), 50th (Base) and 90th (High) percentiles.

```{r}
#| label: tbl-turnaround-time-range
#| tbl-cap: Turnaround time ranges, in minutes, in 2022

read_xlsx(here::here("data", "turnaround_time.xlsx"),
          sheet = "ranges") %>% 
  gt() %>% 
  cols_label(
    category = "Aircraft category"
  ) %>% 
  tab_row_group(
    label = "Scheduled",
    rows = 1:2
  ) %>% 
  tab_row_group(
    label = "Actual",
    rows = 3:4
  ) %>%
  tab_options(
    row_group.background.color = "#F5F7F9",
    row_group.font.weight = "bold"
  ) %>% 
  tab_options(
    table.width = pct(100)
   ) %>% 
   cols_width(
    everything() ~ pct(25),
  ) %>% 
  tab_options(
    column_labels.font.weight = "bold"
  ) %>% 
  tab_source_note(
    source_note = md("*Source: EUROCONTROL – Computed from data supplied by the airline members to CODA*")
  )
```

## Comments

Turnaround time and ground time typically vary as a function of:

* the airport

* the type of flight (short, medium or long-haul)

* the market segment (traditional scheduled airline, low-cost, business aviation, etc.)

* the type of aircraft (B738, A320, etc.)

* the type of service (charter, scheduled, positioning, etc.)

The turnaround process involves activities related to the handling of tasks to 
ensure the cleanliness, safety and efficiency of the next flight. The difference 
between a turnaround and ground time is that an aircraft at its home base airport
will have longer ground time to cover for example for the time it needs for
maintenance. The diagram below shows the scope of the various activities, 
including ground handling time.^[An exhaustive definition and list of the ground handling services is given in [Council Directive 96/67/EC of 15 October 1996](http://eur-lex.europa.eu/legal-content/EN/ALL/?uri=CELEX%3A31996L0067) on access to the ground handling market at Community airports] 

![Overview of activities involved in turnaround process](figures/turnaround_activities.svg){#fig-turnaround-activities}

## Related inputs

@sec-airport-classification [Airport classification](#sec-airport-classification)

## References